### 2016-02-09 HW1/submission/CMW_hw1.tex
Updated the ignore file. Created the tex file for submission.

\documentclass[12pt,a4paper]{article}
\usepackage{amsmath}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage[T1]{fontenc}

\author{
  Cheng, Lidens\\
  \texttt{lidenscheng@email.arizona.edu}
  \and
  McClintock, Tom\\
  \texttt{tmcclintock@email.arizona.edu}
  \and
  Wagoner, Erika\\
  \texttt{wagoner47@email.arizona.edu}
}
\title{Astr 513: Homework 1}

\begin{document}
\maketitle

\section{Random Number Generators}

\section{Designing surveys}

\section{Blackbody Distribution}

\end{document}

### 2016-02-10 HW1/submission/CMW_hw1.tex
Added part 3 to the tex file. Part 3 is now complete.

@@
 \usepackage{graphicx}
+\usepackage{listings}
 \usepackage{fullpage}
 \usepackage[T1]{fontenc}
+
+\lstdefinestyle{custompython}{
+  belowcaptionskip=1\baselineskip,
+  breaklines=true,
+  frame=L,
+  xleftmargin=\parindent,
+  language=bash,
+  basicstyle=\footnotesize\ttfamily,
+  showstringspaces=false,
+  %commentstyle=\itshape\color{purple!40!black},
+  %keywordstyle=\itshape\color{green!40!black},
+  %identifierstyle=\color{blue},
+  %stringstyle=\color{orange},
+}
 
@@
 \section{Blackbody Distribution}
+\subsection{}
+Given our initial distribution:
+\begin{equation*}
+  f(\epsilon;T)d\epsilon = C\frac{\epsilon^2d\epsilon}{\exp{(\epsilon/kT)}-1}
+\end{equation*}
+we first make a change in variables $\epsilon' = \epsilon/kT$. This results
+in the distribution
+\begin{equation*}
+  f(\epsilon')d\epsilon' = A\frac{\epsilon'^2d\epsilon'}{\exp{(\epsilon')}-1}
+\end{equation*}
+where $A=C(kT)^2$.
+\subsection{}
+We now need to find where this function is a maximum numerically. First we
+define our function in Python:
+\begin{lstlisting}[style=custompython]
+  def f(eps):
+      return eps**2/(np.exp(eps)-1)
+\end{lstlisting}
+and find its maximum. This was found to be at the 
+location $\epsilon'\approx1.594$.
+\subsection{}
+We now renormalize by dividing by $f(\epsilon=1.594)\approx0.6476$.
+This yields:
+\begin{equation*}
+  f'(\epsilon')d\epsilon' = B\frac{\epsilon'^2d\epsilon'}{\exp{(\epsilon')}-1}
+\end{equation*}
+where $B=A/0.6476$.
+\subsection{}
+We define a function to return a photon of energy between $\epsilon'_1=0.1$
+and $\epsilon'_2=5.0$.
+\begin{lstlisting}[style=custompython]
+  def en(eps):
+      return np.random.rand()*(5.0-0.1)+0.1
+\end{lstlisting}
+\subsection{}
+We now make repeated draws to this function to get a collection of $N$
+photon energies ${e_i},\ i\in[1,N]$. Additionally, we draw $N$ random
+numbers between $0$ and $1$, ${r_i},\ i\in[1,N]$. If $f'(\epsilon'_i) < r_i$
+then the photon is removed from the list. We do this for $N=100000$.
+\subsection
+Histogramming the photons yields a version of the distribution
+with the area of the bins normalized to $1$.
+\begin{figure*}[!h]
+  \includegraphics[]{HW1_3_6_normalized.png}
+\end{figure*}
+If we simply loop over the energy domain and plot the fraction of accepted
+photons in that bin then we recover $f'$.
+\begin{figure*}[!h]
+  \includegraphics[]{HW1_3_6.png}
+\end{figure*}
 

### 2016-02-11 HW1/submission/CMW_hw1.tex
Problem 1 is done and added to tex file

@@
 \section{Random Number Generators}
+Our pseudo-random number generator (PRNG) of choice is the Mersenne Twister, MT19937.
+It was developed by Makoto Matsumoto and Takuji Nishimura in 1997.
+The algorithm generates a sequence of word vectors of size $w$, which are uniform pseudo-random integers in $[0,2^w-1]$.
+After dividing by $2^w-1$, the word vectors become real numbers in $[0,1]$.
+The linear recurrence relation for each word is a twisted version of a generalized feedback shift register, which consists of bitwise shifts, ORs, XORs, and matrix multiplication [1][2].
+\\\\
+Parameters have been chosen so that MT19937 has a long period of $2^{19937}-1$.
+The output is equidistributed in $k=623$ dimensions for 32-bit words.
+This means it would only exhibit patterns when $k \geq 623$ dimensions [1][2].
+Thus, clustering does not seem to be a big issue with MT19937.\\\\
+MT19937 is not particularly space efficient.
+The generator state takes up 2500 bytes of RAM ($625 \times 32$ bits) [3].
+It passes most of the statistical randomness tests, but not all of the most stringent ones.
+This is most applicable to instances where two seeds are extremely similar.
+In such a case, MT19937 can take a long time before the two sequences of numbers would diverge.
+This issue has been addressed in an update [4].
+In terms of speed, MT19937 has been evaluated to generate approximately $138.889$ million random numbers per second, $7.2$ns for each number, and at $44\%$ of the speed relative to the fastest PRNG [5].
+\\\\
+Although it might not be the most efficient PRNG, MT19937 is extremely portable.
+It comes default in numerous general purpose, programming languages and softwares such as C++, Python, R, MATLAB, and IDL.
+In addition, it is available in a 64-bit variety.
+MT19937 is a reliable PRNG independent of the compiler or system being used.
 
@@
 then the photon is removed from the list. We do this for $N=100000$.
-\subsection
+\subsection{}
 Histogramming the photons yields a version of the distribution
@@
 \end{figure*}
+\newpage
 If we simply loop over the energy domain and plot the fraction of accepted
@@
 \end{figure*}
+\newpage
+
+\section*{References}
+\begin{enumerate}
+  \item   Matsumoto, Makoto, and Takuji Nishimura.
+"Mersenne Twister: A 623-dimensionally Equidistributed Uniform Pseudo-random Number Generator." Transactions on Modeling and Computer Simulation.
+TOMACS ACM Transactions on Modeling and Computer Simulation 8.1 (1998): 3-30.
+  \item Tian, Xiang, and Khaled Benkrid. "Mersenne Twister Random Number Generation on FPGA, CPU and GPU." 2009 NASA/ESA Conference on Adaptive Hardware and Systems (2009): Web.
+  \item Matsumoto, Makoto. "Mersenne Twister with Improved Initialization." Hiroshima University. Web. http://www.math.sci.hiroshima-u.ac.jp/~m-mat/MT/MT2002/emt19937ar.html.
+  \item "Pseudo-Random Number Generator." Boost C++ Libraries. Web. \\ http://www.boost.org/doc/libs/1461/doc/html/boostrandom/reference.html.
+  \item "Performance." Boost C++ Libraries. Web. \\ http://www.boost.org/doc/libs/1460/doc/html/boostrandom/performance.html.
+
+\end{enumerate}
 

### 2016-02-11 HW1/submission/CMW_hw1_v2.tex
Added problem 2 to the tex file.
I can’t compile this on my own to make sure it compiles correctly,
because I’m missing some of the packages apparently. If it doesn’t
compile properly, let me know.

--- 2016-02-11 HW1/submission/CMW_hw1.tex
@@
 \section{Designing surveys}
+We are given a prior probability from the rate of detecting planets of $1-2 R_{\oplus}$ around solar type stars from Kepler, and are told to assume this rate has been measured to high accuracy as being 10\%.
+We are tasked with determining the number of solar type stars we would need to observe to find at least 30 $1-2 R_{\oplus}$ planets in a similar survey observing a different area of the sky with a 90\% and 99\% likelihood.
+To do this, we will want to make of the binomial distribution,
+\begin{equation*}
+  P(n, \rho | N) = \left(N \choose n\right) \rho^n \left(1 - \rho\right)^{N - n},
+\end{equation*}
+where $n$ is the number of successes, $\rho$ is the prior likelihood of obtaining a success, and $N$ is the number of independent observations.
+We will want to solve for $N$ given a probability of 0.9 or 0.99, but there is a modification that needs to be made because we want the probability not of finding exactly 30 of theses planets, but 30 or more of them.
+There are two ways to calculate this probability, either by adding the probabilities of finding $n \in {[30, \infty]}$ planets, or by adding the probabilities of finding $n \in {[0, 29)}$ planets and subtracting this sum from 1.
+Computationally, the latter of these is the most logical, so our probability will actually be
+  \begin{equation*}
+    P(n \ge 30, \rho | N) = \sum_{m = 0}^{29} \left(N \choose m\right) \rho^m \left(1 - \rho\right)^{N - m},
+  \end{equation*}
+  and when solving we will use the substitution that $\rho = 0.1$ and $P(n \ge 30, \rho | N) = 0.9$ or $0.99$.
+
+  Using a false position method root finding algorithm in Python, we were able to find the values of $N$ that would satisfy the above equation for our desired likelihoods. These values are
+  \begin{enumerate}[a)]
+  \item $N = 368$ for a 90\% likelihood
+  \item $N = 435$ for a 99\% likelihood
+  \end{enumerate}
 

### 2016-02-12 HW1/submission/CMW_hw1.tex
Merged into one tex file, compiled and converted to PDF for submission

@@
 \section{Designing surveys}
+We are given a prior probability from the rate of detecting planets of $1-2 R_{\oplus}$ around solar type stars from Kepler, and are told to assume this rate has been measured to high accuracy as being 10\%.
+We are tasked with determining the number of solar type stars we would need to observe to find at least 30 $1-2 R_{\oplus}$ planets in a similar survey observing a different area of the sky with a 90\% and 99\% likelihood.
+To do this, we will want to make of the binomial distribution,
+\begin{equation*}
+  P(n, \rho | N) = \left(N \choose n\right) \rho^n \left(1 - \rho\right)^{N - n},
+\end{equation*}
+where $n$ is the number of successes, $\rho$ is the prior likelihood of obtaining a success, and $N$ is the number of independent observations.
+We will want to solve for $N$ given a probability of 0.9 or 0.99, but there is a modification that needs to be made because we want the probability not of finding exactly 30 of theses planets, but 30 or more of them.
+There are two ways to calculate this probability, either by adding the probabilities of finding $n \in {[30, \infty]}$ planets, or by adding the probabilities of finding $n \in {[0, 29)}$ planets and subtracting this sum from 1.
+Computationally, the latter of these is the most logical, so our probability will actually be
+  \begin{equation*}
+    P(n \ge 30, \rho | N) = \sum_{m = 0}^{29} \left(N \choose m\right) \rho^m \left(1 - \rho\right)^{N - m},
+  \end{equation*}
+  and when solving we will use the substitution that $\rho = 0.1$ and $P(n \ge 30, \rho | N) = 0.9$ or $0.99$.\\\\
+Using a false position method root finding algorithm in Python, we were able to find the values of $N$ that would satisfy the above equation for our desired likelihoods. These values are
+\begin{description}
+  \item [\normalfont a)]$N = 368$ for a 90\% likelihood
+  \item [\normalfont b)]$N = 435$ for a 99\% likelihood
+\end{description}
 
@@
 \newpage
+\noindent
 If we simply loop over the energy domain and plot the fraction of accepted
@@
 \end{figure*}
+
 \newpage
@@
 \section*{References}
+
 \begin{enumerate}
@@
   \item Matsumoto, Makoto. "Mersenne Twister with Improved Initialization." Hiroshima University. Web. http://www.math.sci.hiroshima-u.ac.jp/~m-mat/MT/MT2002/emt19937ar.html.
-  \item "Pseudo-Random Number Generator." Boost C++ Libraries. Web. \\ http://www.boost.org/doc/libs/1461/doc/html/boostrandom/reference.html.
-  \item "Performance." Boost C++ Libraries. Web. \\ http://www.boost.org/doc/libs/1460/doc/html/boostrandom/performance.html.
-
+  \item "Pseudo-Random Number Generator." Boost C++ Libraries. Web. \\ http://www.boost.org/doc/libs/1{\_}46{\_}1/doc/html/boost{\_}random/reference.html.
+  \item "Performance." Boost C++ Libraries. Web. \\ http://www.boost.org/doc/libs/1{\_}46{\_}0/doc/html/boost{\_}random/performance.html.   
 \end{enumerate}

### 2016-03-05 HW2/HW2.tex
Created tex file for HW2

--- 2016-02-10 HW1/submission/CMW_hw1.tex
@@
 }
-\title{Astr 513: Homework 1}
+\title{Astr 513: Homework 2}
 
@@
 
-\section{Random Number Generators}
+\section{Parts A and B}
 
-\section{Designing surveys}
+\section{Part C}
 
-\section{Blackbody Distribution}
-\subsection{}
-Given our initial distribution:
-\begin{equation*}
-  f(\epsilon;T)d\epsilon = C\frac{\epsilon^2d\epsilon}{\exp{(\epsilon/kT)}-1}
-\end{equation*}
-we first make a change in variables $\epsilon' = \epsilon/kT$. This results
-in the distribution
-\begin{equation*}
-  f(\epsilon')d\epsilon' = A\frac{\epsilon'^2d\epsilon'}{\exp{(\epsilon')}-1}
-\end{equation*}
-where $A=C(kT)^2$.
-\subsection{}
-We now need to find where this function is a maximum numerically. First we
-define our function in Python:
-\begin{lstlisting}[style=custompython]
-  def f(eps):
-      return eps**2/(np.exp(eps)-1)
-\end{lstlisting}
-and find its maximum. This was found to be at the 
-location $\epsilon'\approx1.594$.
-\subsection{}
-We now renormalize by dividing by $f(\epsilon=1.594)\approx0.6476$.
-This yields:
-\begin{equation*}
-  f'(\epsilon')d\epsilon' = B\frac{\epsilon'^2d\epsilon'}{\exp{(\epsilon')}-1}
-\end{equation*}
-where $B=A/0.6476$.
-\subsection{}
-We define a function to return a photon of energy between $\epsilon'_1=0.1$
-and $\epsilon'_2=5.0$.
-\begin{lstlisting}[style=custompython]
-  def en(eps):
-      return np.random.rand()*(5.0-0.1)+0.1
-\end{lstlisting}
-\subsection{}
-We now make repeated draws to this function to get a collection of $N$
-photon energies ${e_i},\ i\in[1,N]$. Additionally, we draw $N$ random
-numbers between $0$ and $1$, ${r_i},\ i\in[1,N]$. If $f'(\epsilon'_i) < r_i$
-then the photon is removed from the list. We do this for $N=100000$.
-\subsection
-Histogramming the photons yields a version of the distribution
-with the area of the bins normalized to $1$.
-\begin{figure*}[!h]
-  \includegraphics[]{HW1_3_6_normalized.png}
-\end{figure*}
-If we simply loop over the energy domain and plot the fraction of accepted
-photons in that bin then we recover $f'$.
-\begin{figure*}[!h]
-  \includegraphics[]{HW1_3_6.png}
-\end{figure*}
+\section{Parts D and E}
+
+\newpage
+
+\section*{References}
 

### 2016-03-08 HW2/HW2.tex
Added parts a and b to LaTeX file.
For some reason, when I compile it, mine and Tom’s names are on the
same line, so I’m not sure what’s happening there. If you see any typos
or other errors, let me know.

@@
 \usepackage[T1]{fontenc}
+\usepackage{enumerate}
+\usepackage[makeroom]{cancel}
+\usepackage{hyperref}
 
@@
 
+\begin{enumerate}[a)]
+ \item We can use three kinematic equations to find $u_x$ and $u_y$ in terms of $R$, $g$, and $h$:
+ \begin{enumerate}[(1)]
+  \item $0 = u_y^2 - 2 g h \Rightarrow \boxed{u_y = \sqrt{2 g h}}$
+  \item $0 = u_y - \frac{g t}{2} = \sqrt{2 g h} - \frac{g t}{2} \Rightarrow t = \sqrt{\frac{8 h}{g}}$
+  \item $R = u_x t = u_x \sqrt{\frac{8 h}{6}} \Rightarrow \boxed{u_x = R \sqrt{\frac{g}{8h}}}$
+ \end{enumerate}
+ In the error propagation method, the expected values for $u_x$ and $u_y$ are just the values $u_x^0$ and $u_y^0$ we obtain when we insert the mean values $R_0$, $g_0$, and $h_0$ into the expressions above:
+ \begin{subequations}
+  \begin{align*}
+   u_x^0 &= R_0 \sqrt{\frac{g}{8 h}} \\
+   &= 10.0 \mathrm{m} \sqrt{\frac{9.81 \cancel{\mathrm{m}} \cdot \mathrm{s}^{-2}}{8 \cdot 1.0 \cancel{\mathrm{m}}}}
+  \end{align*}
+  \begin{equation*}
+   \boxed{u_x^0 \approx 11.1 \mathrm{m} \cdot \mathrm{s}^{-1}}
+  \end{equation*}
+ \end{subequations}
+ \begin{subequations}
+  \begin{align*}
+   u_y^0 &= \sqrt{2 g_0 h_0} \\
+   &= \sqrt{2 \cdot 9.81 \mathrm{m} \cdot \mathrm{s}^{-2} \cdot 1.0 \mathrm{m}}
+  \end{align*}
+  \begin{equation*}
+   \boxed{u_y^0 \approx 4.43 \mathrm{m} \cdot \mathrm{s}^{-1}}
+  \end{equation*}
+ \end{subequations}
+
+ When using propagation of errors, the fractional error on some function $f\left(\vec{x}\right)$ with mean $\vec{x}$ values $\vec{x}^0$ and errors $\sigma_i$ can be found by $\frac{\sigma_f}{f} = \sqrt{\displaystyle \sum_{i = 1}^n \left(\frac{\partial \ln f}{\partial x_i}\right)_{\vec{x}^0} \sigma_i^2 }$.
+So we first take the natural log of $u_x$ and $u_y$:
+ \begin{align*}
+  \ln u_x &= \ln R + \frac{1}{2} \ln g - \frac{1}{2} \ln h - \frac{3}{2} \ln 2 \\
+  \ln u_y &= \ln 2 + \ln g + \ln h
+ \end{align*}
+ Then we find that
+ \begin{subequations}
+  \begin{align*}
+   \frac{\sigma_{u_x}}{u_x} &= \sqrt{\left(\frac{\sigma_R}{R_0}\right)^2 + \frac{1}{4} \left(\frac{\sigma_g}{g_0}\right)^2 + \frac{1}{4} \left(\frac{\sigma_h}{h_0}\right)^2} \\
+   &= \sqrt{\left(\frac{0.2\cancel{\mathrm{m}}}{10.0\cancel{\mathrm{m}}}\right)^2 + \frac{1}{4} \left(\frac{0.05 \cancel{\mathrm{m}} \cdot \cancel{\mathrm{s}^{-2}}}{9.81 \cancel{\mathrm{m}} \cdot \cancel{\mathrm{s}^{-2}}}\right)^2 + \frac{1}{4} \left(\frac{0.2 \cancel{\mathrm{m}}}{1.0 \cancel{\mathrm{m}}}\right)^2}
+  \end{align*}
+  \begin{equation*}
+   \boxed{\frac{\sigma_{u_x}}{u_x} \approx 0.10}
+  \end{equation*}
+ \end{subequations}
+ for $\sigma_{u_x}$, and for $\sigma_{u_y}$,
+ \begin{subequations}
+  \begin{align*}
+   \frac{\sigma_{u_y}}{u_y} &= \sqrt{\frac{1}{4} \left(\frac{\sigma_g}{g_0}\right)^2 + \frac{1}{4} \left(\frac{\sigma_h}{h_0}\right)^2} \\
+   &= \frac{1}{2} \sqrt{\left(\frac{0.05 \cancel{\mathrm{m}} \cdot \cancel{\mathrm{s}^{-2}}}{9.81 \cancel{\mathrm{m}} \cdot \cancel{\mathrm{s}^{-2}}}\right)^2 + \left(\frac{0.2 \cancel{\mathrm{m}}}{1.0 \cancel{\mathrm{m}}}\right)^2}
+  \end{align*}
+  \begin{equation*}
+   \boxed{\frac{\sigma_{u_y}}{u_y} \approx 0.10}
+  \end{equation*}
+ \end{subequations}
+
+ \item Now we use Monte Carlo realizations of $R$, $g$, and $h$ assuming Gaussian uncertainties to find the distributions of $u_x$ and $u_y$.
+Using this method, we find that $u_x \approx 11.1\substack{+1.32 \\ -0.99}$ and $u_y \approx 4.43\substack{+0.42 \\ -0.46}$.
+So the means for $u_x$ and $u_y$ seem to agree with the error propagation method, but not the errors.
+The marginalized distributions for $u_x$ and $u_y$, as well as the distribution of $u_y$ vs $u_x$, can be seen in the corner plot in Fig ~\ref{fig:2.1b}.
+
+ \begin{figure}[ht]
+  \centering
+  \includegraphics[keepaspectratio]{hw2_b_corner.pdf}
+  \caption{The distribution of $u_y$ vs $u_x$ as well as the marginalized distributions for $u_x$ and $u_y$.
+The red line in the marginalized distributions shows the answer from the error propagation method, and the red point with error bars is the expected value and error from the error propagation method.}
+  \label{fig:2.1b}
+ \end{figure}
+\end{enumerate}
+
 \section{Part C}

### 2016-03-09 HW2/HW2.tex
Started writing my sections (d,e). Changed equation* to equaton so that we can reference things if we want to.

@@
   \end{align*}
-  \begin{equation*}
+  \begin{equation}
    \boxed{u_x^0 \approx 11.1 \mathrm{m} \cdot \mathrm{s}^{-1}}
-  \end{equation*}
+  \end{equation}
  \end{subequations}
@@
   \end{align*}
-  \begin{equation*}
+  \begin{equation}
    \boxed{u_y^0 \approx 4.43 \mathrm{m} \cdot \mathrm{s}^{-1}}
-  \end{equation*}
+  \end{equation}
  \end{subequations}
@@
   \end{align*}
-  \begin{equation*}
+  \begin{equation}
    \boxed{\frac{\sigma_{u_x}}{u_x} \approx 0.10}
-  \end{equation*}
+  \end{equation}
  \end{subequations}
@@
   \end{align*}
-  \begin{equation*}
+  \begin{equation}
    \boxed{\frac{\sigma_{u_y}}{u_y} \approx 0.10}
-  \end{equation*}
+  \end{equation}
  \end{subequations}
@@
 
-\section{Parts D and E}
+\section{Part D}
+Bayes' Theorem says
+\begin{equation}
+  P(A|B) = \frac{P(B|A)P(A)}{P(B)}
+\end{equation}
+where $P(X|Y)$ is a conditional probability of $X$ given $Y$, and $P(Z)$ is 
+refered to as a \textit{prior} belief on what the value of $Z$ should be.
+Frequentists equate the two conditional probabilities, but this is incorrect
+in the Bayesian approach precisely because you wish to be informed by the 
+priors. In this part, we assume \textit{uniform} priors on our posterior
+parameters $u_x$, $u_y$ and $g$, and in addition to this there is no Jacobian 
+transformation. Thus we can write for the conditional probability
+\begin{equation}
+
+\end{equation}
 

### 2016-03-09 HW2/HW2.tex
Fixed a typo and fixed some of part b.
I noticed a typo in part d, so I went ahead and fixed that. I haven’t
compiled yet because of the empty equation environment towards the end,
but I didn’t change any syntax, so it should still work.

@@
 Using this method, we find that $u_x \approx 11.1\substack{+1.32 \\ -0.99}$ and $u_y \approx 4.43\substack{+0.42 \\ -0.46}$.
-So the means for $u_x$ and $u_y$ seem to agree with the error propagation method, but not the errors.
+So the means and errors for $u_x$ and $u_y$ seem to agree with the error propagation method.
 The marginalized distributions for $u_x$ and $u_y$, as well as the distribution of $u_y$ vs $u_x$, can be seen in the corner plot in Fig ~\ref{fig:2.1b}.
@@
 where $P(X|Y)$ is a conditional probability of $X$ given $Y$, and $P(Z)$ is 
-refered to as a \textit{prior} belief on what the value of $Z$ should be.
+referred to as a \textit{prior} belief on what the value of $Z$ should be.
 Frequentists equate the two conditional probabilities, but this is incorrect

### 2016-03-10 HW2/HW2.tex
Added sections D and E. Added my figures. Put utilities.py in this folder so that you don't need to fiddle with the path. Created a new ipynb.

--- 2016-03-09 HW2/HW2.tex
@@
 So the means for $u_x$ and $u_y$ seem to agree with the error propagation method, but not the errors.
-The marginalized distributions for $u_x$ and $u_y$, as well as the distribution of $u_y$ vs $u_x$, can be seen in the corner plot in Fig ~\ref{fig:2.1b}.
+The marginalized distributions for $u_x$ and $u_y$, as well as the distribution of $u_y$ vs $u_x$, can be seen in the corner plot in \autoref{fig:corner}.
 
@@
 The red line in the marginalized distributions shows the answer from the error propagation method, and the red point with error bars is the expected value and error from the error propagation method.}
-  \label{fig:2.1b}
+  \label{fig:corner}
  \end{figure}
@@
 \section{Part C}
+\label{sec:partC}
 
 \section{Part D}
+\label{sec:partD}
 Bayes' Theorem says
@@
 transformation. Thus we can write for the conditional probability
+\begin{align}
+  \label{eq:bayesian_conditional}
+  P_B(u_x,u_y,g|R,h,g) = & \frac{1}{(2\pi)^{3/2}\sigma_R\sigma_h\sigma_g} \exp\left(-\frac{(2u_xu_y/g - R_0)^2}{2\sigma_R^2}\right)\times \\ &\exp\left(-\frac{(u_y^2/2g-h_0)^2}{2\sigma_h^2}\right) \exp\left(-\frac{(g-g_0)^2}{2\sigma_g^2}\right)
+\end{align}
+where we have written $R$ and $h$ as functions of $u_x$, $u_y$ and $g$. Just as in class,
+this expression is missing the Jacobian term,
+$2u_y^2/g^2$, as calculated in \autoref{sec:partC}. Thus, if we assume flat priors
 \begin{equation}
+  \label{eq:priors_partD}
+  P_{pr}(u_x) = P_{pr}(u_y) = 1
+\end{equation}
+then the ratio of the two probabilities is simply this Jacobian
+\begin{equation}
+  \label{eq:prob_ratio}
+  \frac{P_f}{P_b} = \frac{2u_y^2}{g^2}.
+\end{equation}
+Interestingly, this ratio is not unitless, 
+but this is not an issue since in the Bayesian 
+framework we neglect the normalization terms anyway, 
+which carry units with them.
 
-\end{equation}
+We can marginalize over any two parameters to acquire the third.
+See \autoref{sec:partE} for figures showing these distributions 
+with no priors.
+
+\section{Part E}
+We can also introduce logarithmic priors
+\begin{equation*}
+  \label{eq:log_prior}
+  P_{pr}(X) \propto x^{-1}
+\end{equation*}
+on any of our parameters. The effect is to shift (sometimes slightly)
+the distributions down slightly. In the following three figures
+we plot the marginalized probability for each parameter with and 
+without a logarithmic prior. These probabilities were generated
+by performing a numerical integral over each of the marginalized
+parameters.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[keepaspectratio]{ux_bayesian.pdf}
+  \caption{The marginalized probability of $u_X$. The marginalization
+    over the other two parameters was done numerically.}
+  \label{fig:ux_bayesian}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[keepaspectratio]{uy_bayesian.pdf}
+  \caption{The marginalized probability of $u_y$. The marginalization
+    over the other two parameters was done numerically, however
+    one may perform the marginalization over $u_x$ analytically in
+    this exercise.}
+  \label{fig:uy_bayesian}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[keepaspectratio]{g_bayesian.pdf}
+  \caption{The marginalized probability of $g$. The marginalization
+    over the other two parameters was done numerically, however
+    one may perform the marginalization over $u_x$ analytically in
+    this exercise.}
+  \label{fig:g_bayesian}
+\end{figure}
 

### 2016-03-10 HW2/HW2.tex
Added part c graphs, finished tex file, final pdf

@@
 \label{sec:partC}
+The analytic approach to this problem involves a Jacobian transformation from $(R,h,g)$ to $(u_x,u_y,g)$. To do so, we invert 
+\[ h=\frac{u_y^2}{2g} \ \ \textrm{and} 
+\ \ R=\frac{2u_xu_y}{g} \] to give 
+\[ u_x= \sqrt{2gh} \ \ \textrm{and} 
+\ \ u_y=\sqrt{\frac{g}{8h}} .\]
+Of course, $g=g$. The probability distribution is 
+\begin{equation}
+  P(R,h,g)dRdhdg=P(u_x,u_y,g)du_xdu_ydg.
+\end{equation}
+It can be written in terms of three Gaussian functions 
+\begin{equation}
+  P(R,h,g)=G(R,R_0,\sigma_R)G(h,h_0,\sigma_h)G(g,g_0,\sigma_g).
+\end{equation}
+The Jacobian transformation of the probability distribution is 
+\begin{equation}
+  P(u_x,u_y,g) = G(R,R_0,\sigma_R)G(h,h_0,\sigma_h)G(g,g_0,\sigma_g) J\bigg(\frac{R,h,g}{u_x,u_y,g}\bigg)
+\end{equation}
+where $J$ is the Jacobian. In this transformation from 3 variables to 3 inferences, 
+\begin{equation}J=
+\begin{vmatrix}
+  \frac{\partial R}{\partial u_x} & \frac{\partial R}{\partial u_y} & \frac{\partial R}{\partial g}\\
+  \frac{\partial h}{\partial u_x} & \frac{\partial h}{\partial u_y} & \frac{\partial h}{\partial g}\\
+  \frac{\partial g}{\partial u_x} & \frac{\partial g}{\partial u_y} & \frac{\partial g}{\partial g}\\
+\end{vmatrix}= \frac{2u_y^2}{g^2}. 
+\end{equation}
+With the Jacobian, 
+\begin{eqnarray}\nonumber
+P(u_x,u_y,g)&=& \frac{u_y^2}{\sqrt{2}\pi^{3/2} \sigma_R \sigma_h \sigma_g g^2}\textnormal{exp}\bigg[{\frac {-(\frac{2u_yu_x}{g}-R_0)^2}{2 {\sigma_R} ^2}}\bigg]\\
+&& \textnormal{exp}\bigg[{\frac {-(\frac{u_y^2}{2g}-h_0)^2}{2 {\sigma_h} ^2}}\bigg]\textnormal{exp}\bigg[{\frac {-(g-g_0)^2}{2 {\sigma_g} ^2}}\bigg].
+\end{eqnarray}
+In order to obtain the probability distribution of $u_y$, we need to marginalize Equation 9 over $u_x$ and $g$. Thus, 
+\begin{eqnarray}\nonumber
+P(u_y)&=& \frac{u_y^2}{\sqrt{2}\pi^{3/2} \sigma_R \sigma_h \sigma_g}\textnormal{exp}\bigg[{\frac {-(\frac{u_y^2}{2g}-h_0)^2}{2 {\sigma_h} ^2}}\bigg]\\
+&&\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\textnormal{exp}\bigg[{\frac {-(\frac{2u_yu_x}{g}-R_0)^2}{2 {\sigma_R} ^2}}\bigg]\textnormal{exp}\bigg[{\frac {-(g-g_0)^2}{2 {\sigma_g} ^2}}\bigg]\frac{1}{g^2}du_xdg.
+\end{eqnarray}
+In order to obtain the probability distribution of $u_x$, we need to marginalize Equation 9 over $u_y$ and $g$. Thus, 
+\begin{eqnarray}\nonumber
+P(u_x)&=& \frac{1}{\sqrt{2}\pi^{3/2}\sigma_R \sigma_h \sigma_g} \int_{-\infty}^{\infty}\int_{-\infty}^{\infty} \textnormal{exp}\bigg[{\frac {-(\frac{u_y^2}{2g}-h_0)^2}{2 {\sigma_h} ^2}}\bigg]\\
+&& \textnormal{exp}\bigg[{\frac {-(\frac{2u_yu_x}{g}-R_0)^2}{2 {\sigma_R} ^2}}\bigg]\ \textnormal{exp}\bigg[{\frac {-(g-g_0)^2}{2 {\sigma_g} ^2}}\bigg]\frac{u_y^2}{g^2}du_ydg. 
+\end{eqnarray}
+As seen in \autoref{fig:ux_analytic} and \autoref{fig:uy_analytic}, the shapes of the analytic probability distributions match the shapes of the probability distributions from Monte Carlo simulations.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[keepaspectratio]{ux_analytic.pdf}
+  \caption{The analytic, marginalized probability of $u_x$. The marginalization
+    over the other two parameters was done numerically.}
+  \label{fig:ux_analytic}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[keepaspectratio]{uy_analytic.pdf}
+  \caption{The analytic, marginalized probability of $u_y$. The marginalization
+    over the other two parameters was done numerically.}
+  \label{fig:uy_analytic}
+\end{figure}
 
@@
 We can marginalize over any two parameters to acquire the third.
-See \autoref{sec:partE} for figures showing these distributions 
+See \autoref{fig:ux_bayesian}\autoref{fig:uy_bayesian} for figures showing these distributions 
 with no priors.
@@
 \section{Part E}
+\label{sec:partE}
 We can also introduce logarithmic priors
@@
 
-\newpage
-
-\section*{References}
-
 \end{document}

### 2016-03-29 HW3/HW3.tex
Added file for HW3 and finished problem 1.
Let me know if you see any typos or anything else I should fix/change!

\documentclass[12pt,a4paper]{article}
\usepackage{amsmath}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{fullpage}
\usepackage[T1]{fontenc}
\usepackage{enumerate}
\usepackage[makeroom]{cancel}
\usepackage{hyperref}

\lstdefinestyle{custompython}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  frame=L,
  xleftmargin=\parindent,
  language=bash,
  basicstyle=\footnotesize\ttfamily,
  showstringspaces=false,
  %commentstyle=\itshape\color{purple!40!black},
  %keywordstyle=\itshape\color{green!40!black},
  %identifierstyle=\color{blue},
  %stringstyle=\color{orange},
}

\author{
  Cheng, Lidens\\
  \texttt{lidenscheng@email.arizona.edu}
  \and
  McClintock, Tom\\
  \texttt{tmcclintock@email.arizona.edu}
  \and
  Wagoner, Erika\\
  \texttt{wagoner47@email.arizona.edu}
}
\title{Astr 513: Homework 3}

\begin{document}
\maketitle

\section{Linear models of data with correlated uncertainties}
The definition for $\Delta_i$ will not change when $\rho_i \neq 0$, as it does not depend on the uncertainties. So plugging in for $\vec{u}^\top$, $\vec{Z_i}$, and $\cos \theta$, we have
\begin{equation}
 \label{eqn:Delta}
 \Delta_i = \frac{1}{\sqrt{1 + b^2}} \left(\begin{bmatrix}-b & 1\end{bmatrix} \begin{bmatrix}x_i \\ y_i\end{bmatrix} - a\right),
\end{equation}
and simplifying and squaring gives
\begin{equation}
 \label{eqn:Delta2}
 \Delta_i^2 = \frac{\left(y_i - a - b x_i\right)^2}{1 + b^2}.
\end{equation}

However, $\Sigma_i^2$ is defined with the covariance matrix, which will change when the uncertainties are correlated. We have that $\Sigma_i^2 = \vec{u}^\top\, \mathbf{S_i}\, \vec{u}$, and the covariance matrix for correlated uncertainties with $\sigma_{xyi} = \rho_i \sigma_{xi} \sigma_{yi} \neq 0$ is given as, for the data point $(x_i, y_i)$,
\begin{equation}
 \label{eqn:Si}
 \mathbf{S_i} = \begin{bmatrix}\sigma_{xi}^2 & \rho_i \sigma_{xi} \sigma_{yi} \\ \rho_i \sigma_{xi} \sigma_{yi} & \sigma_{yi}^2\end{bmatrix}.
\end{equation}
Thus, we find
\begin{align}
 \Sigma_i^2 &= \frac{1}{1 + b^2} \begin{bmatrix}-b & 1\end{bmatrix} \begin{bmatrix}\sigma_{xi}^2 & \rho_i \sigma_{xi} \sigma_{yi} \\ \rho_i \sigma_{xi} \sigma_{yi} & \sigma_{yi}^2\end{bmatrix} \begin{bmatrix}-b \\ 1\end{bmatrix} \nonumber \\
 &= \frac{1}{1 + b^2} \begin{bmatrix}\sigma_{xi} \left(-b \sigma_{xi} + \rho_i \sigma_{yi}\right) & \sigma_{yi} \left(-b \rho_i \sigma_{xi} + \sigma_{yi}\right)\end{bmatrix} \begin{bmatrix}-b \\ 1\end{bmatrix} \nonumber \\
 &= \frac{1}{1 + b^2} \left(b^2 \sigma_{xi}^2 - b \rho_i \sigma_{xi} \sigma_{yi} - b \rho_i \sigma_{xi} \sigma_{yi} + \sigma_{yi}^2\right) \nonumber \\
 \Rightarrow \Sigma_i^2 &= \frac{\sigma_{yi}^2 -2 b \rho_i \sigma_{xi} \sigma_{yi} + b^2 \sigma_{xi}^2}{1 + b^2}. \label{eqn:Sigma2}
\end{align}

When we then take $\Delta_i^2/2 \Sigma_i^2$, we will see the factor of $1 + b^2$ cancel from both terms, and are left with
\begin{equation}
 \label{eqn:ans1}
 \boxed{\frac{\Delta_i^2}{2 \Sigma_i^2} = \frac{\left(y_i - a - b x_i\right)^2}{2 \left(\sigma_{yi}^2 - 2 b \rho_i \sigma_{xi} \sigma_{yi} + b^2 \sigma_{xi}^2\right)}}
\end{equation}

\section{The Tully-Fisher relation}

\end{document}

### 2016-04-04 HW3/utilities.py
Added an ipython notebook and utilities in case we need it.

@@ -1,0 +1,77 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import corner
+
+"""
+Given two flat distributions in x1 and x2,
+return two distributions z1 and z2 that are Gaussian
+distributed.
+"""
+def boxmuller(x1,x2):
+    z1 = np.sqrt(-2*np.log(x1))*np.cos(2*np.pi*x2)
+    z2 = np.sqrt(-2*np.log(x1))*np.sin(2*np.pi*x2)
+    return z1,z2
+
+"""
+This 
+"""
+def get_fake_gaussian_data(N):
+    x1,x2 = np.random.rand(N),np.random.rand(N)
+    return boxmuller(x1,x2)
+
+"""
+Given a histogram P with bin centers y,
+find the levels of P (Pjs) that contain some fraction of the total
+probability under them (Cjs). Then, find the values of P and y (P0 , y0)
+that contain C0 amount of probability within them.
+"""
+def conf_intervals(y,P,C0=0.683,k=1000):
+    #Step 1 find where the maximum value is
+    i_Pmax = np.argmax(P)
+    Pmax, y_Pmax, M  = P[i_Pmax],y[i_Pmax], len(y)
+    dy = (np.max(y)-np.min(y))/(M-1)
+    
+    #Step 2 grid P with K Points, j = 1...k
+    #Step 3 find C_j, the integral over P(y)H[P(y)-Pj], i.e. the fraction
+    #in the grid of P
+    Pjs = []
+    Cjs = []
+    for j in range(1,k+1):
+        Pj = (j-1.0)/(k-1.0)*Pmax
+        Pjs.append(Pj)
+        Cj = 0
+        for i in range(2,M):
+            if P[i] > Pj:
+                Cj+=dy*P[i]
+            else:
+                Cj+=0
+        Cjs.append(Cj)
+
+    #Step 4 scan through the Cjs and find j
+    j = 0
+    for i in range(len(Cjs)-1):
+        if Cjs[i+1] < C0:
+            j = i
+            break
+    
+    #Step 5 interpolate for P at that point
+    P0 = (Pjs[j+1]-Pjs[j])/(Cjs[j+1]-Cjs[j]) * (C0 - Cjs[j]) + Pjs[j]
+    
+    #Step 6 interpolate again to find i_left, i_right and y0_left and y0_right
+    i_left,iIright = 0,0
+    for l in range(len(P)-1):
[…]

### 2016-04-04 HW3/HW3.tex
Updated question 2. Finished the ipython notebook.

@@
 
+ \begin{figure}[ht]
+  \centering
+  \includegraphics[keepaspectratio]{figures/tully_fisher.eps}
+  \caption{The magnitude and angular velocity form a power law.}
+  \label{fig:tully_fisher}
+ \end{figure}
+
+We took our catalog and maximized the likelihood given in \autoref{eqn:ans1}.
+Our raw data is shown in \autoref{fig:tully_fisher}, and the best
+fit is shown in \autoref{fig:tully_fisher_fit}.
+
+ \begin{figure}[ht]
+  \centering
+  \includegraphics[keepaspectratio]{figures/tully_fisher_fit.eps}
+  \caption{The magnitude and angular velocity form a power law.
+  The best fit Tully-Fisher relation is overplotted with the
+  points from the galaxy catalog.}
+  \label{fig:tully_fisher_fit}
+ \end{figure}
+
 \end{document}

### 2016-04-11 Final_Project/ms.tex
Making the initial tex file for the final project.

--- 2016-03-10 HW2/HW2.tex
@@
 }
-\title{Astr 513: Homework 2}
+\title{Astr 513: Final Project}
 
@@
 
-\section{Parts A and B}
-
-\begin{enumerate}[a)]
- \item We can use three kinematic equations to find $u_x$ and $u_y$ in terms of $R$, $g$, and $h$:
- \begin{enumerate}[(1)]
-  \item $0 = u_y^2 - 2 g h \Rightarrow \boxed{u_y = \sqrt{2 g h}}$
-  \item $0 = u_y - \frac{g t}{2} = \sqrt{2 g h} - \frac{g t}{2} \Rightarrow t = \sqrt{\frac{8 h}{g}}$
-  \item $R = u_x t = u_x \sqrt{\frac{8 h}{6}} \Rightarrow \boxed{u_x = R \sqrt{\frac{g}{8h}}}$
- \end{enumerate}
- In the error propagation method, the expected values for $u_x$ and $u_y$ are just the values $u_x^0$ and $u_y^0$ we obtain when we insert the mean values $R_0$, $g_0$, and $h_0$ into the expressions above:
- \begin{subequations}
-  \begin{align*}
-   u_x^0 &= R_0 \sqrt{\frac{g}{8 h}} \\
-   &= 10.0 \mathrm{m} \sqrt{\frac{9.81 \cancel{\mathrm{m}} \cdot \mathrm{s}^{-2}}{8 \cdot 1.0 \cancel{\mathrm{m}}}}
-  \end{align*}
-  \begin{equation}
-   \boxed{u_x^0 \approx 11.1 \mathrm{m} \cdot \mathrm{s}^{-1}}
-  \end{equation}
- \end{subequations}
- \begin{subequations}
-  \begin{align*}
-   u_y^0 &= \sqrt{2 g_0 h_0} \\
-   &= \sqrt{2 \cdot 9.81 \mathrm{m} \cdot \mathrm{s}^{-2} \cdot 1.0 \mathrm{m}}
-  \end{align*}
-  \begin{equation}
-   \boxed{u_y^0 \approx 4.43 \mathrm{m} \cdot \mathrm{s}^{-1}}
-  \end{equation}
- \end{subequations}
-
- When using propagation of errors, the fractional error on some function $f\left(\vec{x}\right)$ with mean $\vec{x}$ values $\vec{x}^0$ and errors $\sigma_i$ can be found by $\frac{\sigma_f}{f} = \sqrt{\displaystyle \sum_{i = 1}^n \left(\frac{\partial \ln f}{\partial x_i}\right)_{\vec{x}^0} \sigma_i^2 }$.
-So we first take the natural log of $u_x$ and $u_y$:
- \begin{align*}
-  \ln u_x &= \ln R + \frac{1}{2} \ln g - \frac{1}{2} \ln h - \frac{3}{2} \ln 2 \\
-  \ln u_y &= \ln 2 + \ln g + \ln h
- \end{align*}
- Then we find that
- \begin{subequations}
-  \begin{align*}
-   \frac{\sigma_{u_x}}{u_x} &= \sqrt{\left(\frac{\sigma_R}{R_0}\right)^2 + \frac{1}{4} \left(\frac{\sigma_g}{g_0}\right)^2 + \frac{1}{4} \left(\frac{\sigma_h}{h_0}\right)^2} \\
-   &= \sqrt{\left(\frac{0.2\cancel{\mathrm{m}}}{10.0\cancel{\mathrm{m}}}\right)^2 + \frac{1}{4} \left(\frac{0.05 \cancel{\mathrm{m}} \cdot \cancel{\mathrm{s}^{-2}}}{9.81 \cancel{\mathrm{m}} \cdot \cancel{\mathrm{s}^{-2}}}\right)^2 + \frac{1}{4} \left(\frac{0.2 \cancel{\mathrm{m}}}{1.0 \cancel{\mathrm{m}}}\right)^2}
-  \end{align*}
-  \begin{equation}
-   \boxed{\frac{\sigma_{u_x}}{u_x} \approx 0.10}
-  \end{equation}
- \end{subequations}
- for $\sigma_{u_x}$, and for $\sigma_{u_y}$,
- \begin{subequations}
-  \begin{align*}
-   \frac{\sigma_{u_y}}{u_y} &= \sqrt{\frac{1}{4} \left(\frac{\sigma_g}{g_0}\right)^2 + \frac{1}{4} \left(\frac{\sigma_h}{h_0}\right)^2} \\
-   &= \frac{1}{2} \sqrt{\left(\frac{0.05 \cancel{\mathrm{m}} \cdot \cancel{\mathrm{s}^{-2}}}{9.81 \cancel{\mathrm{m}} \cdot \cancel{\mathrm{s}^{-2}}}\right)^2 + \left(\frac{0.2 \cancel{\mathrm{m}}}{1.0 \cancel{\mathrm{m}}}\right)^2}
-  \end{align*}
-  \begin{equation}
-   \boxed{\frac{\sigma_{u_y}}{u_y} \approx 0.10}
-  \end{equation}
- \end{subequations}
-
- \item Now we use Monte Carlo realizations of $R$, $g$, and $h$ assuming Gaussian uncertainties to find the distributions of $u_x$ and $u_y$.
-Using this method, we find that $u_x \approx 11.1\substack{+1.32 \\ -0.99}$ and $u_y \approx 4.43\substack{+0.42 \\ -0.46}$.
-So the means for $u_x$ and $u_y$ seem to agree with the error propagation method, but not the errors.
-The marginalized distributions for $u_x$ and $u_y$, as well as the distribution of $u_y$ vs $u_x$, can be seen in the corner plot in \autoref{fig:corner}.
-
- \begin{figure}[ht]
-  \centering
-  \includegraphics[keepaspectratio]{hw2_b_corner.pdf}
-  \caption{The distribution of $u_y$ vs $u_x$ as well as the marginalized distributions for $u_x$ and $u_y$.
-The red line in the marginalized distributions shows the answer from the error propagation method, and the red point with error bars is the expected value and error from the error propagation method.}
-  \label{fig:corner}
- \end{figure}
-\end{enumerate}
-
-\section{Part C}
-\label{sec:partC}
-
-\section{Part D}
-\label{sec:partD}
-Bayes' Theorem says
-\begin{equation}
-  P(A|B) = \frac{P(B|A)P(A)}{P(B)}
-\end{equation}
-where $P(X|Y)$ is a conditional probability of $X$ given $Y$, and $P(Z)$ is 
-refered to as a \textit{prior} belief on what the value of $Z$ should be.
-Frequentists equate the two conditional probabilities, but this is incorrect
-in the Bayesian approach precisely because you wish to be informed by the 
-priors. In this part, we assume \textit{uniform} priors on our posterior
-parameters $u_x$, $u_y$ and $g$, and in addition to this there is no Jacobian 
-transformation. Thus we can write for the conditional probability
-\begin{align}
-  \label{eq:bayesian_conditional}
-  P_B(u_x,u_y,g|R,h,g) = & \frac{1}{(2\pi)^{3/2}\sigma_R\sigma_h\sigma_g} \exp\left(-\frac{(2u_xu_y/g - R_0)^2}{2\sigma_R^2}\right)\times \\ &\exp\left(-\frac{(u_y^2/2g-h_0)^2}{2\sigma_h^2}\right) \exp\left(-\frac{(g-g_0)^2}{2\sigma_g^2}\right)
-\end{align}
-where we have written $R$ and $h$ as functions of $u_x$, $u_y$ and $g$. Just as in class,
-this expression is missing the Jacobian term,
-$2u_y^2/g^2$, as calculated in \autoref{sec:partC}. Thus, if we assume flat priors
-\begin{equation}
-  \label{eq:priors_partD}
-  P_{pr}(u_x) = P_{pr}(u_y) = 1
-\end{equation}
-then the ratio of the two probabilities is simply this Jacobian
-\begin{equation}
-  \label{eq:prob_ratio}
-  \frac{P_f}{P_b} = \frac{2u_y^2}{g^2}.
-\end{equation}
-Interestingly, this ratio is not unitless, 
-but this is not an issue since in the Bayesian 
-framework we neglect the normalization terms anyway, 
-which carry units with them.
-
-We can marginalize over any two parameters to acquire the third.
-See \autoref{sec:partE} for figures showing these distributions 
-with no priors.
-
-\section{Part E}
-We can also introduce logarithmic priors
-\begin{equation*}
-  \label{eq:log_prior}
-  P_{pr}(X) \propto x^{-1}
-\end{equation*}
-on any of our parameters. The effect is to shift (sometimes slightly)
-the distributions down slightly. In the following three figures
-we plot the marginalized probability for each parameter with and 
-without a logarithmic prior. These probabilities were generated
-by performing a numerical integral over each of the marginalized
-parameters.
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[keepaspectratio]{ux_bayesian.pdf}
-  \caption{The marginalized probability of $u_X$. The marginalization
-    over the other two parameters was done numerically.}
-  \label{fig:ux_bayesian}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[keepaspectratio]{uy_bayesian.pdf}
-  \caption{The marginalized probability of $u_y$. The marginalization
-    over the other two parameters was done numerically, however
-    one may perform the marginalization over $u_x$ analytically in
-    this exercise.}
-  \label{fig:uy_bayesian}
-\end{figure}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[keepaspectratio]{g_bayesian.pdf}
-  \caption{The marginalized probability of $g$. The marginalization
-    over the other two parameters was done numerically, however
-    one may perform the marginalization over $u_x$ analytically in
-    this exercise.}
-  \label{fig:g_bayesian}
-\end{figure}
-
-\newpage
+\section{Type Ia Supernovae}
 
@@
 
+\begin{enumerate}
+  \item test
+\end{enumerate}
+
 \end{document}

### 2016-04-25 Final_Project/cosmosis_code/run_script.py
Added a script that will hold the call to our code, as well as the files that will make up the files we use for cosmosis.

@@ -1,0 +1,25 @@
+"""
+This is the script used to run the code for the final project.
+"""
+
+import os
+
+do_analysis = True
+do_postprocess = False
+
+if do_analysis:
+    datapath = "/path/to/the/data"
+    covpath  = "/path/to/the/cov/matrix"
+    
+    cosmology = "final_cosmology.ini"
+    pipeline = "final_pipeline.ini"
+    
+    command = "datapath=%s covpath=%s cosmology=%s cosmosis %s"\
+              %(datapath,covpath,cosmology,pipeline)
+
+"""
+Now that the code is run, we can do post processing below.
+"""
+
+if do_postprocess:
+    x=0 #garbage for now

### 2016-04-28 Final_Project/ms.tex
Began writing document
I began writing on the data. I also added a bibliography file that we
can edit and the apj.bst for the bibliography style I put in the tex
document, although if you guys would prefer a different style, we can
change that.

@@
 \usepackage{hyperref}
+\usepackage{natbib}
+\bibliographystyle{apj}
 
@@
 }
+
+% Roman numerals command
+\makeatletter
+\newcommand*{\rom}[1]{\expandafter\@slowromancap\romannumeral #1@}
+\makeatother
+
+% Some definitions for writing SN, SNe, SN Ia, and SNe Ia...
+\newcommand{\sn}{\mbox{SN}}
+\newcommand{\sne}{\mbox{SNe}}
+\newcommand{\sna}{\mbox{SN \rom{1}a}}
+\newcommand{\snea}{\mbox{SNe \rom{1}a}}
 
@@
 
-\section*{References}
+\section{Data}
+\label{sec:data}
 
-\begin{enumerate}
-  \item test
-\end{enumerate}
+Our data is compiled from \citet{betoule2014} and \citet{rest2014} blah blah blah...
+
+\subsection{The JLA sample}
+\label{sec:betoule}
+
+\citet{betoule2014} define their ``joint light-curve analysis'' (JLA) sample, a total of 740 type Ia supernovae (SNe Ia) %If we define the abbreviations elsewhere, this can change -- ELW
+ compiled from a combination of data from the Sloan Digital Sky Survey (SDSS), the Supernova Legacy Survey (SNLS), the Hubble Space Telescope (HST), and several other experiments.
+The 613 SDSS \snea{} come from the SDSS supernova survey (SDSS-SNS) of SDSS-II \citep{sako2014} with redshifts of $0.05 < z < 0.4$, and 374 of them are selected from those that were confirmed in the spectroscopic follow-up.
+The rest of the \sne{} come from the compilation by \citet{conley2011}, which includes 242 spectroscopically confirmed \snea{} from SNLS at redshifts $0.2 < z < 1$, 14 \snea{} from HST at very high redshifts of $0.7 < z < 1.4$, and 123 low redshift ($z < 0.08$) from a variety of sources.
+We made use of the publicly available distance moduli, redshifts, and covariance matrix provided by \citet{betoule2014}\footnote{Data is available for download via \url{http://cdsweb.u-strasbrg.fr/cgi-bin/qcat?J/A+A/568/A22}}.
+We note that the distance moduli are binned, so that they are only computed and given on a grid of redshifts.
+The distance modulus can then be well approximated by a piecewise linear function of $\log(z)$ that is given in a single redshift bin $z_b \le z < z_{b+1}$ as \citep[see][Appendix E.1]{betoule2014}
+\begin{equation}
+\label{eq:binnedMu}
+\bar{\mu}(z) = (1 - \alpha) \mu_b + \alpha \mu_{b+1},
+\end{equation}
+where they have defined $\alpha = \log\left(\frac{z}{z_b}\right)/\log\left(\frac{z_{b+1}}{z_b}\right)$ and $\mu_i$ is the distance modulus at control point $z_i$.
+Therefore, the data and covariance matrix are merely given at the 31 control points log-spaced in the redshift range $0.01 < z < 1.3$.
+
+%\section*{References}
+%
+%\begin{enumerate}
+%  \item test
+%\end{enumerate}
+
+\bibliography{final}
 

### 2016-04-28 Final_Project/cosmosis_code/astr_513_SNIA.py
Edited git ignore to ignore the manuscript pdf. Wrote more code.

@@ -1,0 +1,60 @@
+"""
+This is the module that compares the supernovae data
+to our model and returns a likelihood.
+
+"""
+
+from cosmosis.datablock import names as section_names
+from cosmosis.datablock import option_section
+from cosmosis.runtime.declare import declare_module
+import numpy as np
+
+cosmo = section_names.cosmological_parameters
+likes = section_names.likelihoods
+
+"""
+The setup function runs only once at the start
+"""
+def setup(options):
+
+    #Open the data and covariance matrix
+    #Then calculate the inverse covariance matrix
+    datapath = options[option_section,"datapath"]
+    covpath = options[option_section,"covpath"]
+    data = np.genfromtxt(datapath)
+    cov = np.genfromtxt(covpath)
+    icov = np.linalg.inv(cov)
+    z,mu = data.T
+    
+    #Pass it all forward
+    return (z,mu,cov,icov)
+
+"""
+This function evaluates the likelihood.
+"""
+def execute(block,config):
+    z,mu,cov,icov = config #Passed from setup
+
+    #Read in the cosmology
+    h = block[cosmo,"h0"]
+    om = block[cosmo,"omega_m"]
+    ode = 1.0 - om #Omega_dark_energy
+
+    #Create a model for mu(z)
+    #TODO: decide how to implement this
+    mu_model = z*0 #MODEL GOES HERE
+
+    #Calculate the likelihood
+    LL = 0
+    for i in range(len(cov)):
+        for j in range(len(cov[i])):
+            LL += -0.5 * (mu[i]-mu_model[i])*icov[i,j]*(mu[j]-mu_model[j])
+    return LL
+    
+"""
+This function has to be here for design reasons.
+It is only really used if the module is written in
+C/C++/Fortran.
+"""
+def cleanup(self):
+    return 0

### 2016-04-28 Final_Project/ms.tex
Added todo notes package to the manuscript. Added linebreaks so it looks better in emacs.

@@
 }
+
+\usepackage{todonotes}
+\newcommand\TM[1]{\todo[color=green!40, inline, size=\small]{TM: #1}}
 
@@
 \section{Type Ia Supernovae}
+\TM{TODO: Complete this section.}
 
@@
 \label{sec:data}
-
+\TM{TODO: Complete this section.}
 Our data is compiled from \citet{betoule2014} and \citet{rest2014} blah blah blah...
@@
 
-\citet{betoule2014} define their ``joint light-curve analysis'' (JLA) sample, a total of 740 type Ia supernovae (SNe Ia) %If we define the abbreviations elsewhere, this can change -- ELW
- compiled from a combination of data from the Sloan Digital Sky Survey (SDSS), the Supernova Legacy Survey (SNLS), the Hubble Space Telescope (HST), and several other experiments.
-The 613 SDSS \snea{} come from the SDSS supernova survey (SDSS-SNS) of SDSS-II \citep{sako2014} with redshifts of $0.05 < z < 0.4$, and 374 of them are selected from those that were confirmed in the spectroscopic follow-up.
-The rest of the \sne{} come from the compilation by \citet{conley2011}, which includes 242 spectroscopically confirmed \snea{} from SNLS at redshifts $0.2 < z < 1$, 14 \snea{} from HST at very high redshifts of $0.7 < z < 1.4$, and 123 low redshift ($z < 0.08$) from a variety of sources.
-We made use of the publicly available distance moduli, redshifts, and covariance matrix provided by \citet{betoule2014}\footnote{Data is available for download via \url{http://cdsweb.u-strasbrg.fr/cgi-bin/qcat?J/A+A/568/A22}}.
-We note that the distance moduli are binned, so that they are only computed and given on a grid of redshifts.
-The distance modulus can then be well approximated by a piecewise linear function of $\log(z)$ that is given in a single redshift bin $z_b \le z < z_{b+1}$ as \citep[see][Appendix E.1]{betoule2014}
+\citet{betoule2014} define their ``joint light-curve analysis'' (JLA) 
+sample, a total of 740 type Ia supernovae (SNe Ia) 
+%If we define the abbreviations elsewhere, this can change -- ELW
+compiled from a combination of data from the Sloan Digital Sky 
+Survey (SDSS), the Supernova Legacy Survey (SNLS), the Hubble Space 
+Telescope (HST), and several other experiments. The 613 SDSS \snea{} 
+come from the SDSS supernova survey (SDSS-SNS) of SDSS-II \citep{sako2014} 
+with redshifts of $0.05 < z < 0.4$, and 374 of them are selected from those 
+that were confirmed in the spectroscopic follow-up. The rest of the 
+\sne{} come from the compilation by \citet{conley2011}, which includes 
+242 spectroscopically confirmed \snea{} from SNLS at redshifts $0.2 < z < 1$, 
+14 \snea{} from HST at very high redshifts of $0.7 < z < 1.4$, and 123 
+low redshift ($z < 0.08$) from a variety of sources. We made use of the 
+publicly available distance moduli, redshifts, and covariance 
+matrix provided by \citet{betoule2014}\footnote{Data is available 
+for download via \url{http://cdsweb.u-strasbrg.fr/cgi-bin/qcat?J/A+A/568/A22}}. 
+We note that the distance moduli are binned, so that they are only computed and 
+given on a grid of redshifts. The distance modulus can then be well 
+approximated by a piecewise linear function of $\log(z)$ that is given 
+in a single redshift bin $z_b \le z < z_{b+1}$ as 
+\citep[see][Appendix E.1]{betoule2014}
 \begin{equation}
@@
 \end{equation}
-where they have defined $\alpha = \log\left(\frac{z}{z_b}\right)/\log\left(\frac{z_{b+1}}{z_b}\right)$ and $\mu_i$ is the distance modulus at control point $z_i$.
-Therefore, the data and covariance matrix are merely given at the 31 control points log-spaced in the redshift range $0.01 < z < 1.3$.
+where they have defined 
+$\alpha = \log\left(\frac{z}{z_b}\right)/\log\left(\frac{z_{b+1}}{z_b}\right)$ 
+and $\mu_i$ is the distance modulus at control point $z_i$. 
+Therefore, the data and covariance matrix are merely given at the 
+31 control points log-spaced in the redshift range $0.01 < z < 1.3$.
 

### 2016-04-28 Final_Project/ms.tex
Added todo notes commands for each of us to write messages.
Added an example figure and added new sections.

@@
 \newcommand\TM[1]{\todo[color=green!40, inline, size=\small]{TM: #1}}
+\newcommand\EW[1]{\todo[color=blue!40, inline, size=\small]{TM: #1}}
+\newcommand\LC[1]{\todo[color=magenta!40, inline, size=\small]{TM: #1}}
 
@@
 
+\subsection{Covariances}
+\label{sec:covariances}
+\TM{A subsection that explains how the covariance matrix looks. Basically
+  a brief discussion about how it was made and then a picture of the 
+  {\it correlation} matrix. Example figure below.}
+
+\begin{figure}
+ \includegraphics[width=\linewidth]{figures/clusters-randoms_noqc_correlation_subset.eps}
+ \caption{Correlation matrix example. We will just have one panel.}
+ \label{fig:cmatrix}
+\end{figure}
+
+\section{$\mu$ Model}
+\label{sec:model}
+\TM{This explains how to go from a cosmology to $\mu(z)$.}
+
+\subsection{Likelihood}
+\label{sec:likelihood}
+\TM{This has our likelihood.}
+
 %\section*{References}

### 2016-04-28 Final_Project/Betoule2014Data/make_correlation_matrix.py

@@ -1,0 +1,4 @@
+"""
+A short script that takes Betoule's covariance matrix and computs the 
+correlation matrix, and then makes a figure.
+"""

### 2016-04-28 Final_Project/Betoule2014Data/make_correlation_matrix.py
Created the correlation matrix figure and added it to the text.

@@ -2,3 +2,35 @@
 A short script that takes Betoule's covariance matrix and computs the 
 correlation matrix, and then makes a figure.
 """
+import numpy as np
+
+#Pull out the z and mus
+data = np.genfromtxt("tablef1.dat")
+z,mu = data.T
+
+#Get the covariance matrix
+cov  = np.genfromtxt("tablef2.dat")
+
+#Define the correlation matrix
+corr = np.zeros_like(cov)
+for i in range(len(cov)):
+    for j in range(len(cov[i])):
+        corr[i,j] = cov[i,j]/(np.sqrt(cov[i,i]*cov[j,j]))
+
+#Visualize
+import matplotlib.pyplot as plt
+plt.rc('text',usetex=True,fontsize=28)
+fig,ax = plt.subplots()
+cax = ax.imshow(corr,interpolation="nearest",\
+                    vmin=-1,vmax=1,cmap=plt.get_cmap("bwr"),\
+                    extent=[z[0],z[-1],z[-1],z[0]])
+ticks = np.linspace(-1.0,1.0,11)
+cbar = fig.colorbar(cax,ticks=ticks)
+plt.rc
+fs = 28
+ax.set_xlabel("Redshift $z$",fontsize=fs)
+ax.set_ylabel("Redshift $z$",fontsize=fs)
+plt.subplots_adjust(bottom=0.15)
+
+plt.gcf().savefig("correlation.eps")
+plt.show()

### 2016-04-28 Final_Project/ms.tex
Created the correlation matrix figure and added it to the text.
Renamed final.bib to msbib.bib to avoid any possible confusion.

@@
 \TM{A subsection that explains how the covariance matrix looks. Basically
-  a brief discussion about how it was made and then a picture of the 
-  {\it correlation} matrix. Example figure below.}
+  a brief discussion about how it was made.}
 
 \begin{figure}
- \includegraphics[width=\linewidth]{figures/clusters-randoms_noqc_correlation_subset.eps}
- \caption{Correlation matrix example. We will just have one panel.}
+ \includegraphics[width=\linewidth]{figures/correlation.eps}
+ \caption{Correlation matrix of $\mu_0(z)$ from the
+ \citet{betoule2014} sample.}
  \label{fig:cmatrix}
@@
 
-\section{$\mu$ Model}
+\section{Distance Modulus Model}
 \label{sec:model}
@@
 
+\subsection{Results}
+\TM{Discuss the results and show the corner plot. Example plot included.}
+
+\begin{figure}
+ \includegraphics[width=\linewidth]{figures/corner_example.eps}
+ \caption{Example corner plot.}
+ \label{fig:corner}
+\end{figure}
+
 %\section*{References}
@@
 
-\bibliography{final}
+\bibliography{msbib}
 

### 2016-04-29 Final_Project/Rest2014Data/make_correlation_matrix.py
Created new correlation matrices and added them to the paper. Put in section labels for all sections. Added sections about Planck.

@@ -5,11 +5,11 @@ correlation matrix, and then makes a figure.
 import numpy as np
 
 #Pull out the z and mus
-data = np.genfromtxt("tablef1.dat")
+data = np.genfromtxt("redshift_distanceMod.dat")
 z,mu = data.T
 
 #Get the covariance matrix
-cov  = np.genfromtxt("tablef2.dat")
+cov  = np.genfromtxt("covarianceMatrix.dat")
 
 #Define the correlation matrix
 corr = np.zeros_like(cov)
@@ -32,5 +32,5 @@ ax.set_xlabel("Redshift $z$",fontsize=fs)
 ax.set_ylabel("Redshift $z$",fontsize=fs)
 plt.subplots_adjust(bottom=0.15)
 
-plt.gcf().savefig("Betoule_correlation.eps")
+plt.gcf().savefig("Rest_correlation.eps")
 plt.show()

### 2016-04-29 Final_Project/ms.tex
Created new correlation matrices and added them to the paper. Put in section labels for all sections. Added sections about Planck.
Started writing the Planck section.

@@
 \newcommand\LC[1]{\todo[color=magenta!40, inline, size=\small]{TM: #1}}
+\newcommand\Planck{{\it Planck}\ }
 
@@
 \section{Type Ia Supernovae}
+\label{sec:SNIa}
 \TM{TODO: Complete this section.}
+
+\section{Planck 2015 Results}
+\label{sec:planck}
+In 2009 the European Space Agency launched the \Planck satellite in order
+to measure the Cosmis Microwave Background (CMB) with unparalleled precision.
+By measuring the small anisotropies in the temperature of the CMB across
+the entire sky one calculates the angular power spectrum, which can be
+predicted from first principles (see, e.g., \citet{dodelson2003}).
 
@@
 \begin{figure}
- \includegraphics[width=\linewidth]{figures/correlation.eps}
+ \includegraphics[width=\linewidth]{figures/Betoule_correlation.eps}
  \caption{Correlation matrix of $\mu_0(z)$ from the
  \citet{betoule2014} sample.}
+ \label{fig:cmatrix}
+\end{figure}
+
+\begin{figure}
+ \includegraphics[width=\linewidth]{figures/Rest_correlation.eps}
+ \caption{Correlation matrix of $\mu_0(z)$ from the
+ \citet{rest2014} sample.}
  \label{fig:cmatrix}
@@
 
-\subsection{Results}
-\TM{Discuss the results and show the corner plot. Example plot included.}
+\subsection{Planck Priors}
+\label{sec:planck_priors}
+
+\section{Results}
+\label{sec:results}
+\TM{Discuss the results and show the corner plots. Example plot included.
+  We should have one corner plot without priors and one with priors.}
 
@@
 
-%\section*{References}
-%
-%\begin{enumerate}
-%  \item test
-%\end{enumerate}
+\section{Conclusions}
+\label{sec:conclusions}
+\TM{Talk about how the inclusions if priors changes the analysis.
+  State whether the priors are strong or uninformative.}
 

### 2016-04-29 Final_Project/ms.tex
Tentatively done with the planck section. Can definitely go into detail on why we chose this.

@@
 \newcommand\LC[1]{\todo[color=magenta!40, inline, size=\small]{TM: #1}}
+
 \newcommand\Planck{{\it Planck}\ }
+\newcommand\Cosmosis{{\it Cosmosis}\ }
 
@@
 the entire sky one calculates the angular power spectrum, which can be
-predicted from first principles (see, e.g., \citet{dodelson2003}).
+modeled using first principles (see, e.g., \citet{dodelson2003}).
+By using a likelihood method similar to our own, the authors of
+\citet{planck2013} estimated 14 nuisance parameters and 
+six cosmological parameters, including the geometrical parameters,
+$\Omega_M$, $\Omega_\Lambda$ and $\Omega_k$, that we use in our
+SNIa analysis. The only prior used by \citet{planck2013}
+on cosmological parameters was from the WMAP result \citep{wmap2011}
+for the optical depth $\tau$. We do not use this parameter
+in our SNIa model.
+
+\citet{planck2013} report marginalized
+posterior distributions for all parameters,
+which we report in Table~XXX. This prior is implemented
+in the planck2015 module implemented in \Cosmosis \citep{cosmosis}.
 

### 2016-04-30 Final_Project/ms.tex
Updated Section 1 and added some new references

@@
 
-\section{Type Ia Supernovae}
+\section{Type \rom{1}a Supernovae}
 \label{sec:SNIa}
 \TM{TODO: Complete this section.}
+\EW{You guys might want to read this over and tell me if you think I should
+  change/add anything.}
+A type \rom{1}a supernova (\sna{}) is a supernova with very little hydrogen or
+helium absorption. It is believed that the progenitor for these special
+supernovae are white dwarfs that are part of a binary system. These white
+dwarfs accrete mass from their companion stars or merge with a companion white
+dwarf or neutron star, causing them to exceed the Chandrasekhar limit
+($1.44 M_\odot$). When this happens, carbon and oxygen fusion begins in the
+core, and the temperature increases. However, the white dwarf does not expand
+and cool as a main sequence star would, since it is supported by degeneracy
+pressure rather than thermal, so the core experiences a runaway fusion
+reaction, releasing $\sim 10^{44}$ J of energy
+\citep[see, e.g.,][chap. 18.4]{ryden2010}, which is enough to overcome the
+gravitational energy and the star explodes. Because all \snea{} are caused by a
+similar reaction, they all have similar luminosities, and more importantly, the
+light curves all have the same shape: a sharp maximum followed by a smooth,
+gradual decay. Studying nearby \snea{} has shown that the rate of decline of
+the light curve is correlated to the peak luminosity, and this correlation has
+been calibrated using Cephied variable stars located in the same galaxies as
+nearby \sna{}. With this information, we can estimate peak luminosity of an
+\sna{} from its light curve, thereby providing us with a standard candle with
+which to measure distance. They were first used in this way by the
+Cal\'{a}n/Tololo Supernova Survey, and it was using \snea{} that
+\citet{riess1998} and \citet{perlmutter1999} were able to discover the
+accelerating expansion rate of the Universe for which three of the authors
+later won the Nobel Prize in physics.
 
@@
 \citet{betoule2014} define their ``joint light-curve analysis'' (JLA) 
-sample, a total of 740 type Ia supernovae (SNe Ia) 
-%If we define the abbreviations elsewhere, this can change -- ELW
-compiled from a combination of data from the Sloan Digital Sky 
+sample, a total of 740 \snea{} compiled from a combination of data from the Sloan Digital Sky 
 Survey (SDSS), the Supernova Legacy Survey (SNLS), the Hubble Space 

### 2016-05-01 Final_Project/ms.tex
Wrote things for Sections 3 and 4

@@
 \newcommand\TM[1]{\todo[color=green!40, inline, size=\small]{TM: #1}}
-\newcommand\EW[1]{\todo[color=blue!40, inline, size=\small]{TM: #1}}
-\newcommand\LC[1]{\todo[color=magenta!40, inline, size=\small]{TM: #1}}
+\newcommand\EW[1]{\todo[color=blue!40, inline, size=\small]{EW: #1}}
+\newcommand\LC[1]{\todo[color=magenta!40, inline, size=\small]{LC: #1}}
 
@@
 \label{sec:SNIa}
-\TM{TODO: Complete this section.}
 \EW{You guys might want to read this over and tell me if you think I should
@@
 $\Omega_M$, $\Omega_\Lambda$ and $\Omega_k$, that we use in our
-SNIa analysis. The only prior used by \citet{planck2013}
+{\sna} analysis. The only prior used by \citet{planck2013}
 on cosmological parameters was from the WMAP result \citep{wmap2011}
 for the optical depth $\tau$. We do not use this parameter
-in our SNIa model.
+in our {\sna} model.
 
@@
 \label{sec:data}
-\TM{TODO: Complete this section.}
-Our data is compiled from \citet{betoule2014} and \citet{rest2014} blah blah blah...
+Our {\sna} data is compiled from the samples analyzed in \citet{betoule2014} and \citet{rest2014}.
+In the following subsections, we will specify the surveys and detail the composition of the two samples.
+Additionally, we will describe the publicly released data from the two papers.
 
@@
 matrix provided by \citet{betoule2014}\footnote{Data is available 
-for download via \url{http://cdsweb.u-strasbrg.fr/cgi-bin/qcat?J/A+A/568/A22}}. 
+for download via \url{http://cdsarc.u-strasbg.fr/viz-bin/qcat?J/A+A/568/A22}}. 
 We note that the distance moduli are binned, so that they are only computed and 
@@
 
+\subsection{The Pan-STARRS1 sample}
+\label{sec:rest}
+
+\citet{rest2014} uses the first 1.5 year of the first telescope of the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS1) and a compilation of low-redshift {\sna} studies.
+The low-redshift sample consists of {\snea} collected by Calan/Tololo (Hamuy et al.
+1996a), CfA1 (Riess et al.
+1999), CfA2 (Jha et al.
+2006), CfA3 (Hicken et al.
+2009a), CSP (Contreras et al.
+2010), and CfA4 (Hicken et al.
+2012).
+After three quality cuts, 113 spectroscopically confirmed {\snea} from Pan-STARRS1 ($0.03 < z < 0.65$) and 222 light curves from 197 low-redshift {\snea} are used for cosmological analysis.
+
+The publicly available data from \citet{rest2014} contains all light curves, light curve fit information, covariance matrix, and each {\sn} host information.\footnote{Data can be accessed at \url{http://ipp0012.ifa.hawaii.edu/SN1aLC/}}.
+For our analysis, we utilized the redshifts, distance moduli, and covariance matrix of the final sample of 310 {{\snea} or 335 light curves.
+
 \subsection{Covariances}
@@
   a brief discussion about how it was made.}
+\LC{Wrote about Rest's cov matrix. Still need description of Betoule's.}
 
@@
 \end{figure}
+
+The publicly released covariance matrix from \citet{rest2014} is comprised of systematic errors for the final sample.
+These systematics, defined in \citet{scolnic2014}, include flux calibration errors, selection effects, light curve fitting, host galaxy relations, Milky Way extinction corrections, and coherent velocity flows.
+Calculating the covariance matrix involves varying a systematic parameter and determining the difference between the original distance and the new distance.
+The covariance between {\sn} $i$ and {\sn} $j$ is
+\begin{equation}
+\label{eq:RestCovMatrix}
+C_{ij}=\displaystyle \sum_{k = 1}^K \left(\frac{\partial \mu_i}{\partial S_k}\right)\left(\frac{\partial \mu_j}{\partial S_k}\right)\left(\sigma_{S_k}\right)^2,
+\end{equation}
+where the summation is over $K$ systematics, $\sigma_{S_k}$ is each systematic error magnitude, and $\partial \mu$ is the difference in distance modulus from varying a systematic parameter.
 
@@
 \TM{This explains how to go from a cosmology to $\mu(z)$.}
+\LC{Need more here but I wrote most of the relevant eqns}
+The main objective of \citet{betoule2014} and \citet{rest2014} is to place constraints on cosmological parameters $\Omega_M$, $\Omega_\Lambda$, and $w$.
+In order to so, we must examine the relationship between distance modulus and cosmology.
+The distance modulus $\mu$ is determined from the luminosity distance $d_L$ by
+\begin{equation}
+\label{eq:mu_dL}
+\mu=5 \text{log}(d_L)+25. 
+\end{equation}
+The luminosity distance depends on the cosmology by 
+\begin{equation}
+\label{eq:luminosityDistance}
+d_L(z; w, \Omega_M, \Omega_\Lambda, H_0)= (1+z)|\Omega_k|^{-1/2}\mathcal{S}_{k}\displaystyle\left[c|\Omega_k|^{1/2}\int_{0}^{z}\frac{dz'}{H(z')}\right],
+\end{equation}
+where $\Omega_k \equiv 1-\Omega_M-\Omega_{\Lambda}$. The function $\mathcal{S}_{k}$ depends on the curvature density $\Omega_k$ in that 
+\[
+    \mathcal{S}_{k}= 
+\begin{cases}
+    \text{sin}(x),&  \Omega_k < 0\\
+    x,& \Omega_k = 0\\
+    \text{sinh}(x),& \Omega_k > 0
+\end{cases}
+\]
+and the function $H(z)$ is 
+\begin{equation}
+\label{eq:hubbleFunction}
+H(z)= H_0\displaystyle\left[\Omega_M(1+z)^3+\Omega_{\Lambda}(1+z)^{3(1+w)}+\Omega_k(1+z)^2\right]^{1/2}.
+\end{equation}
+Three (?) cosmologies are of interest to us: \LC{Not sure which ones we're interested in. $\Lambda CDM$, $\Omega_M=1$, ?}
 

### 2016-05-02 Final_Project/ms.tex
Added linebreaks and removed some of my comments. Added a new comment in section 4.0.

@@
 \label{sec:data}
-Our {\sna} data is compiled from the samples analyzed in \citet{betoule2014} and \citet{rest2014}.
-In the following subsections, we will specify the surveys and detail the composition of the two samples.
+Our {\sna} data is compiled from the samples analyzed in 
+\citet{betoule2014} and \citet{rest2014}. In the following subsections, 
+we will specify the surveys and detail the composition of the two samples. 
 Additionally, we will describe the publicly released data from the two papers.
@@
 \citet{betoule2014} define their ``joint light-curve analysis'' (JLA) 
-sample, a total of 740 \snea{} compiled from a combination of data from the Sloan Digital Sky 
+sample, a total of 740 \snea{} compiled from a combination of 
+data from the Sloan Digital Sky 
 Survey (SDSS), the Supernova Legacy Survey (SNLS), the Hubble Space 
@@
 matrix provided by \citet{betoule2014}\footnote{Data is available 
-for download via \url{http://cdsarc.u-strasbg.fr/viz-bin/qcat?J/A+A/568/A22}}. 
+for download via 
+\url{http://cdsarc.u-strasbg.fr/viz-bin/qcat?J/A+A/568/A22}}. 
 We note that the distance moduli are binned, so that they are only computed and 
@@
 
-\citet{rest2014} uses the first 1.5 year of the first telescope of the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS1) and a compilation of low-redshift {\sna} studies.
-The low-redshift sample consists of {\snea} collected by Calan/Tololo (Hamuy et al.
-1996a), CfA1 (Riess et al.
-1999), CfA2 (Jha et al.
-2006), CfA3 (Hicken et al.
-2009a), CSP (Contreras et al.
-2010), and CfA4 (Hicken et al.
-2012).
-After three quality cuts, 113 spectroscopically confirmed {\snea} from Pan-STARRS1 ($0.03 < z < 0.65$) and 222 light curves from 197 low-redshift {\snea} are used for cosmological analysis.
+\citet{rest2014} uses the first 1.5 year of the first telescope of 
+the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS1) 
+and a compilation of low-redshift {\sna} studies. The low-redshift 
+sample consists of {\snea} collected by Calan/Tololo (Hamuy et al. 1996a), 
+CfA1 (Riess et al. 1999), CfA2 (Jha et al. 2006), CfA3 (Hicken et al. 2009a), 
+CSP (Contreras et al. 2010), and CfA4 (Hicken et al. 2012). After 
+three quality cuts, 113 spectroscopically confirmed {\snea} from 
+Pan-STARRS1 ($0.03 < z < 0.65$) and 222 light curves from 197 
+low-redshift {\snea} are used for cosmological analysis.
 
-The publicly available data from \citet{rest2014} contains all light curves, light curve fit information, covariance matrix, and each {\sn} host information.\footnote{Data can be accessed at \url{http://ipp0012.ifa.hawaii.edu/SN1aLC/}}.
-For our analysis, we utilized the redshifts, distance moduli, and covariance matrix of the final sample of 310 {{\snea} or 335 light curves.
+The publicly available data from \citet{rest2014} contains all light 
+curves, light curve fit information, covariance matrix, and each {\sn} 
+host information.\footnote{Data can be accessed at 
+\url{http://ipp0012.ifa.hawaii.edu/SN1aLC/}}. For our analysis, we 
+utilized the redshifts, distance moduli, and covariance matrix of the 
+final sample of 310 {{\snea} or 335 light curves.
 
@@
 \label{sec:covariances}
-\TM{A subsection that explains how the covariance matrix looks. Basically
-  a brief discussion about how it was made.}
 \LC{Wrote about Rest's cov matrix. Still need description of Betoule's.}
@@
 
-The publicly released covariance matrix from \citet{rest2014} is comprised of systematic errors for the final sample.
-These systematics, defined in \citet{scolnic2014}, include flux calibration errors, selection effects, light curve fitting, host galaxy relations, Milky Way extinction corrections, and coherent velocity flows.
-Calculating the covariance matrix involves varying a systematic parameter and determining the difference between the original distance and the new distance.
+The publicly released covariance matrix from \citet{rest2014} 
+is comprised of systematic errors for the final sample. These systematics, 
+defined in \citet{scolnic2014}, include flux calibration errors, 
+selection effects, light curve fitting, host galaxy relations, Milky 
+Way extinction corrections, and coherent velocity flows. Calculating 
+the covariance matrix involves varying a systematic parameter and determining 
+the difference between the original distance and the new distance. 
 The covariance between {\sn} $i$ and {\sn} $j$ is 
@@
 \end{equation}
-where the summation is over $K$ systematics, $\sigma_{S_k}$ is each systematic error magnitude, and $\partial \mu$ is the difference in distance modulus from varying a systematic parameter.
+where the summation is over $K$ systematics, $\sigma_{S_k}$ is each 
+systematic error magnitude, and $\partial \mu$ is the difference in 
+distance modulus from varying a systematic parameter.
 
@@
 \label{sec:model}
-\TM{This explains how to go from a cosmology to $\mu(z)$.}
 \LC{Need more here but I wrote most of the relevant eqns}
-The main objective of \citet{betoule2014} and \citet{rest2014} is to place constraints on cosmological parameters $\Omega_M$, $\Omega_\Lambda$, and $w$.
-In order to so, we must examine the relationship between distance modulus and cosmology.
-The distance modulus $\mu$ is determined from the luminosity distance $d_L$ by
+The main objective of \citet{betoule2014} and \citet{rest2014} is 
+to place constraints on cosmological parameters $\Omega_M$, $\Omega_\Lambda$, 
+and $w$. In order to so, we must examine the relationship between distance 
+modulus and cosmology. The distance modulus $\mu$ is determined from 
+the luminosity distance $d_L$ by 
 \begin{equation}
@@
 \end{equation}
-where $\Omega_k \equiv 1-\Omega_M-\Omega_{\Lambda}$. The function $\mathcal{S}_{k}$ depends on the curvature density $\Omega_k$ in that 
+where $\Omega_k \equiv 1-\Omega_M-\Omega_{\Lambda}$. The function 
+$\mathcal{S}_{k}$ depends on the curvature density $\Omega_k$ in that 
 \[
@@
 \end{equation}
-Three (?) cosmologies are of interest to us: \LC{Not sure which ones we're interested in. $\Lambda CDM$, $\Omega_M=1$, ?}
+Three (?) cosmologies are of interest to us: 
+\LC{Not sure which ones we're interested in. $\Lambda CDM$, $\Omega_M=1$, ?}
+\TM{I think there are only two, one with $\Omega_k=0$ and one with it
+  being unconstrained.}
 

### 2016-05-02 Final_Project/ms.tex
Edited SN section, added Betoule cov matrix description.

@@
 \label{sec:SNIa}
-\EW{You guys might want to read this over and tell me if you think I should
-  change/add anything.}
+\EW{Should I cite the Nobel Prize information document where I found out about 
+the stuff in the last few sentences? I'm not entirely sure what that would be considered...}
 A type \rom{1}a supernova (\sna{}) is a supernova with very little hydrogen or
@@
 pressure rather than thermal, so the core experiences a runaway fusion
-reaction, releasing $\sim 10^{44}$ J of energy
-\citep[see, e.g.,][chap. 18.4]{ryden2010}, which is enough to overcome the
-gravitational energy and the star explodes. Because all \snea{} are caused by a
-similar reaction, they all have similar luminosities, and more importantly, the
-light curves all have the same shape: a sharp maximum followed by a smooth,
-gradual decay. Studying nearby \snea{} has shown that the rate of decline of
-the light curve is correlated to the peak luminosity, and this correlation has
-been calibrated using Cephied variable stars located in the same galaxies as
-nearby \sna{}. With this information, we can estimate peak luminosity of an
-\sna{} from its light curve, thereby providing us with a standard candle with
-which to measure distance. They were first used in this way by the
-Cal\'{a}n/Tololo Supernova Survey, and it was using \snea{} that
-\citet{riess1998} and \citet{perlmutter1999} were able to discover the
-accelerating expansion rate of the Universe for which three of the authors
-later won the Nobel Prize in physics.
+reaction, releasing $\sim 10^{44}$ J of energy \citep[see, e.g.,][chap.
+18.4]{ryden2010}, which is enough to overcome the gravitational energy and the star explodes.
+Because all \snea{} are caused by a similar reaction, they all have similar luminosities, and more importantly, the light curves all have the same shape: a sharp maximum followed by a smooth, gradual decay.
+Studying nearby \snea{} has shown that the rate of decline of the light curve is correlated to the peak luminosity, and this correlation has been calibrated using Cephied variable stars located in the same galaxies as nearby \sna{}.
+With this information, we can estimate peak luminosity of an \sna{} from its light curve, thereby providing us with a standard candle with which to measure distance.
+They were first used in this way in the 1990s by two independent teams, the Supernova Cosmology Project and the Cal\'{a}n/Tololo Supernova Survey.
+It was using distances from \snea{} from these two projects that \citet{riess1998} and \citet{perlmutter1999} were able to discover the accelerating expansion rate of the Universe in 1998.
+It was for this work that Saul Perlmutter, Brian P.
+Schmidt, and Adam G.
+Reiss won the Nobel Prize in physics in 2011.
 
@@
 \LC{Wrote about Rest's cov matrix. Still need description of Betoule's.}
+\EW{Wrote about Beoule's covariance matrix.}
 
@@
 
+\citet{betoule2014} choose to first calculate the covariance matrix $\mathbf{C}_{\mathbf{\eta}}$ of the 3 light curve fit-parameters $\mathbf{\eta}$ by adding the individual covariance matrices they get from statistical errors propagated from light-curve fit uncertainties and 7 sources of systematic uncertainties they consider from their photometric calibartion, light-curve model uncertainty, bias correction, a mass step uncertainty introduced when correcting magnitudes due to host mass, Galactic extinction, peculiar velocities, and contamination by non-\sna{}.
+They then propagate these errors to the distance modulus and use three terms to account for uncertainty in cosmological redshift by the equation
+\begin{equation}
+\label{eq:BetouleCovMatrix}
+\mathbf{C} = \mathbf{A} \mathbf{C}_\mathbf{\eta} \mathbf{A}^\dag + \mathrm{diag}\left(\frac{5 \sigma_z}{z \log 10}\right)^2 + \mathrm{diag}\left(\sigma_{\mathrm{lens}}^2\right) + \mathrm{diag}\left(\sigma_{\mathrm{coh}}^2\right),
+\end{equation}
+where $\mathbf{A}$ is the matrix of coefficients in the distance modulus, and redshift uncertainty is accounted for in the first diagonal term due to peculiar velocities, the second due to magnitude variation from gravitational lensing, and the third from intrinsic \sn{} magnitude variation not included in other terms.
+They use $c \sigma_z = 150 \mathrm{km}\, \mathrm{s}^{-1}$, $\sigma_{\mathrm{lens}} = 0.055 \times z$, and a value of $\sigma_{\mathrm{coh}}$ for each survey found using a restricted log-likelihood technique.
+The elements of the covariance matrix available online are calculated on the redshift control points $z_b$ as described in \S \ref{sec:betoule}.
+
 \begin{figure}
@@
 \LC{Need more here but I wrote most of the relevant eqns}
+\EW{Do we want/need citations for the equations here?}
 The main objective of \citet{betoule2014} and \citet{rest2014} is 

### 2016-05-02 Final_Project/cosmosis_code/astr_513_SNIA.py
Almost done with the cosmosis code. Model is incomplete.

@@ -12,6 +12,7 @@ import numpy as np
 cosmo = section_names.cosmological_parameters
 likes = section_names.likelihoods
 
+
 """
 The setup function runs only once at the start
 """
@@ -45,11 +46,13 @@ def execute(block,config):
     mu_model = z*0 #MODEL GOES HERE
 
     #Calculate the likelihood
-    LL = 0
-    for i in range(len(cov)):
-        for j in range(len(cov[i])):
-            LL += -0.5 * (mu[i]-mu_model[i])*icov[i,j]*(mu[j]-mu_model[j])
-    return LL
+    LL = -1.0
+    #for i in range(len(cov)):
+        #for j in range(len(cov[i])):
+            #LL += -0.5 * (mu[i]-mu_model[i])*icov[i,j]*(mu[j]-mu_model[j])
+    block[likes,"SNIA_LIKE"]=LL
+
+    return 0 #Signal that it finished fine
     
 """
 This function has to be here for design reasons.

### 2016-05-02 Final_Project/cosmosis_code/run_script.py

@@ -8,14 +8,19 @@ do_analysis = True
 do_postprocess = False
 
 if do_analysis:
-    datapath = "/path/to/the/data"
-    covpath  = "/path/to/the/cov/matrix"
+    base = "/home/tmcclintock/Desktop/Github_stuff/Astr513-Homeworks/Final_Project/Betoule2014Data/"
+    datapath = base+"tablef1.dat"
+    covpath  = base+"tablef2.dat"
     
-    cosmology = "final_cosmology.ini"
-    pipeline = "final_pipeline.ini"
+    path = "my_tests/astr513/"
+    cosmology = path+"final_cosmology.ini"
+    pipeline = path+"final_pipeline.ini"
+    cosmosis = "mpirun -n 4 cosmosis --mpi"
     
-    command = "datapath=%s covpath=%s cosmology=%s cosmosis %s"\
-              %(datapath,covpath,cosmology,pipeline)
+    command = "datapath=%s covpath=%s cosmology=%s %s %s"\
+              %(datapath,covpath,cosmology,cosmosis,pipeline)
+
+    os.system(command)
 
 """
 Now that the code is run, we can do post processing below.

### 2016-05-02 Final_Project/cosmosis_code/astr_513_SNIA.py
final version of the code.

@@ -8,10 +8,16 @@ from cosmosis.datablock import names as section_names
 from cosmosis.datablock import option_section
 from cosmosis.runtime.declare import declare_module
 import numpy as np
+from scipy import integrate
 
 cosmo = section_names.cosmological_parameters
 likes = section_names.likelihoods
 
+"""
+adot/a \propto E_z
+"""
+def E_z(z,om,ode,ok=0):
+    return np.sqrt(om*(1.0+z)**3+ok*(1.0+z)**2+ode)
 
 """
 The setup function runs only once at the start
@@ -43,13 +49,17 @@ def execute(block,config):
 
     #Create a model for mu(z)
     #TODO: decide how to implement this
-    mu_model = z*0 #MODEL GOES HERE
+    dc = np.ones_like(z)
+    for i in range(len(z)):
+        dc[i] = 3000.0/h * integrate.quad(E_z,0,z[i],args=(om,ode))[0]
+    dl = (1+z)*dc
+    mu_model = 5*np.log10(dl)+25
 
     #Calculate the likelihood
-    LL = -1.0
-    #for i in range(len(cov)):
-        #for j in range(len(cov[i])):
-            #LL += -0.5 * (mu[i]-mu_model[i])*icov[i,j]*(mu[j]-mu_model[j])
+    LL = 0.0
+    for i in range(len(cov)):
+        for j in range(len(cov[i])):
+            LL += -0.5 * (mu[i]-mu_model[i])*icov[i,j]*(mu[j]-mu_model[j])
     block[likes,"SNIA_LIKE"]=LL
 
     return 0 #Signal that it finished fine

### 2016-05-02 Final_Project/cosmosis_code/run_script.py

@@ -15,7 +15,8 @@ if do_analysis:
     path = "my_tests/astr513/"
     cosmology = path+"final_cosmology.ini"
     pipeline = path+"final_pipeline.ini"
-    cosmosis = "mpirun -n 4 cosmosis --mpi"
+    cosmosis = "cosmosis"
+    #cosmosis = "mpirun -n 4 cosmosis --mpi"
     
     command = "datapath=%s covpath=%s cosmology=%s %s %s"\
               %(datapath,covpath,cosmology,cosmosis,pipeline)

### 2016-05-02 Final_Project/ms.tex
Expanded on the correlation matrices. Added some comments.

@@
 the stuff in the last few sentences? I'm not entirely sure what that would be considered...}
+\TM{No, because it is common knowledge. I think though that since this is meant to emulate
+  a research paper that we don't need to point out that it was what the NP was awarded for.}
 A type \rom{1}a supernova (\sna{}) is a supernova with very little hydrogen or
@@
 pressure rather than thermal, so the core experiences a runaway fusion
-reaction, releasing $\sim 10^{44}$ J of energy \citep[see, e.g.,][chap.
-18.4]{ryden2010}, which is enough to overcome the gravitational energy and the star explodes.
-Because all \snea{} are caused by a similar reaction, they all have similar luminosities, and more importantly, the light curves all have the same shape: a sharp maximum followed by a smooth, gradual decay.
-Studying nearby \snea{} has shown that the rate of decline of the light curve is correlated to the peak luminosity, and this correlation has been calibrated using Cephied variable stars located in the same galaxies as nearby \sna{}.
-With this information, we can estimate peak luminosity of an \sna{} from its light curve, thereby providing us with a standard candle with which to measure distance.
-They were first used in this way in the 1990s by two independent teams, the Supernova Cosmology Project and the Cal\'{a}n/Tololo Supernova Survey.
-It was using distances from \snea{} from these two projects that \citet{riess1998} and \citet{perlmutter1999} were able to discover the accelerating expansion rate of the Universe in 1998.
-It was for this work that Saul Perlmutter, Brian P.
-Schmidt, and Adam G.
-Reiss won the Nobel Prize in physics in 2011.
+reaction, releasing $\sim 10^{44}$ J of energy \citep[see, e.g.,][chap. 18.4]{ryden2010}, 
+which is enough to overcome the gravitational energy and the star explodes. Because all 
+\snea{} are caused by a similar reaction, they all have similar luminosities, and more 
+importantly, the light curves all have the same shape: a sharp maximum followed by a smooth, 
+gradual decay. Studying nearby \snea{} has shown that the rate of decline of the light curve 
+is correlated to the peak luminosity, and this correlation has been calibrated using Cephied 
+variable stars located in the same galaxies as nearby \sna{}. With this information, one can 
+estimate peak luminosity of an \sna{} from its light curve, thereby providing a 
+standard candle with which to measure distance. They were first used in this way in the 
+1990s by two independent teams, the Supernova Cosmology Project and the Cal\'{a}n/Tololo 
+Supernova Survey. It was using distances from \snea{} from these two projects that 
+\citet{riess1998} and \citet{perlmutter1999} were able to discover the accelerating expansion 
+rate of the Universe in 1998. 
+%It was for this work that Saul Perlmutter, Brian P. Schmidt, 
+%and Adam G. Reiss won the Nobel Prize in physics in 2011.
+\TM{Changed grammer in this section, since ``we'' did not actually do these things.
+  Review and collapse bubbles.}
 
@@
 \label{sec:betoule}
-
 \citet{betoule2014} define their ``joint light-curve analysis'' (JLA) 
@@
 \label{sec:rest}
-
 \citet{rest2014} uses the first 1.5 year of the first telescope of 
@@
 final sample of 310 {{\snea} or 335 light curves.
+\TM{There is some kind of undefined reference error in the above line. 
+  It doesn't alter the pdf document, but can we figure out what is going on?}
 
@@
 \label{sec:covariances}
-\LC{Wrote about Rest's cov matrix. Still need description of Betoule's.}
-\EW{Wrote about Beoule's covariance matrix.}
-
-\begin{figure}
- \includegraphics[width=\linewidth]{figures/Betoule_correlation.eps}
- \caption{Correlation matrix of $\mu_0(z)$ from the
- \citet{betoule2014} sample.}
- \label{fig:cmatrix}
-\end{figure}
-
 The publicly released covariance matrix from \citet{rest2014} 
@@
 
-\citet{betoule2014} choose to first calculate the covariance matrix $\mathbf{C}_{\mathbf{\eta}}$ of the 3 light curve fit-parameters $\mathbf{\eta}$ by adding the individual covariance matrices they get from statistical errors propagated from light-curve fit uncertainties and 7 sources of systematic uncertainties they consider from their photometric calibartion, light-curve model uncertainty, bias correction, a mass step uncertainty introduced when correcting magnitudes due to host mass, Galactic extinction, peculiar velocities, and contamination by non-\sna{}.
-They then propagate these errors to the distance modulus and use three terms to account for uncertainty in cosmological redshift by the equation
+\citet{betoule2014} choose to first calculate the covariance 
+matrix $\mathbf{C}_{\mathbf{\eta}}$ of the 3 light curve fit-parameters $\mathbf{\eta}$ 
+by adding the individual covariance matrices they get from statistical errors propagated 
+from light-curve fit uncertainties and 7 sources of systematic uncertainties 
+they consider from their photometric calibartion, light-curve model uncertainty, 
+bias correction, a mass step uncertainty introduced when correcting magnitudes due 
+to host mass, Galactic extinction, peculiar velocities, and contamination by 
+non-\sna{}. They then propagate these errors to the distance modulus and use three 
+terms to account for uncertainty in cosmological redshift by the equation
 \begin{equation}
@@
 \end{equation}
-where $\mathbf{A}$ is the matrix of coefficients in the distance modulus, and redshift uncertainty is accounted for in the first diagonal term due to peculiar velocities, the second due to magnitude variation from gravitational lensing, and the third from intrinsic \sn{} magnitude variation not included in other terms.
-They use $c \sigma_z = 150 \mathrm{km}\, \mathrm{s}^{-1}$, $\sigma_{\mathrm{lens}} = 0.055 \times z$, and a value of $\sigma_{\mathrm{coh}}$ for each survey found using a restricted log-likelihood technique.
-The elements of the covariance matrix available online are calculated on the redshift control points $z_b$ as described in \S \ref{sec:betoule}.
+where $\mathbf{A}$ is the matrix of coefficients in the distance modulus, 
+and redshift uncertainty is accounted for in the first diagonal term due to 
+peculiar velocities, the second due to magnitude variation from gravitational lensing, 
+and the third from intrinsic \sn{} magnitude variation not included in other terms. 
+They use $c \sigma_z = 150 \mathrm{km}\, \mathrm{s}^{-1}$, 
+$\sigma_{\mathrm{lens}} = 0.055 \times z$, and a value of $\sigma_{\mathrm{coh}}$ for each 
+survey found using a restricted log-likelihood technique. The elements of the covariance 
+matrix available online are calculated on the redshift control points $z_b$ as 
+described in \S \ref{sec:betoule}.
+
+For both of these data sets, the covariance matrix can be visualized using a related
+quantity known as the correlation matrix $Corr$. When two quantities are (anti)correlated,
+the value of one quantity (decreases)increases when the other quantity increases.
+If the variables one has measured are independent, then the correlation between
+differen variables should be $\approx0$.
+
+A correlation matrix for $N$ variables is $N\times N$ in size, and the $ij$th
+element is the correlation between the $i$th and $j$th variables. Therefore, 
+the correlation matrix for \citet{betoule2014} is $31\times31$, while
+for \citet{rest2014} is $335\times335$.
+
+The correlation matrix is calculated from the covariance matrix according to
+\begin{equation}
+  \label{eq:correlation}
+  Corr_{i,j} = \frac{\mathbf{C}_{i,j}}{\sqrt{\mathbf{C}_{i,i}}\sqrt{\mathbf{C}_{j,j}}}.
+\end{equation}
+Note that the correlation of a variable with itself is 1 by definition.
+The correlation matrices for both of our data sets can be seen below in
+figures \autoref{fig:cmatrix_betoule} and \autoref{cmatrix_rest}.
+
+\TM{TODO: Discuss how the correlation matrices for the two data sets are different
+  and hypothesize on what this might mean.}
+
+\begin{figure}
+ \includegraphics[width=\linewidth]{figures/Betoule_correlation.eps}
+ \caption{Correlation matrix of $\mu_0(z)$ from the
+ \citet{betoule2014} sample.
+ \TM{Need to improve this caption.}}
+ \label{fig:cmatrix_betoule}
+\end{figure}
 
@@
  \caption{Correlation matrix of $\mu_0(z)$ from the
- \citet{rest2014} sample.}
- \label{fig:cmatrix}
+ \citet{rest2014} sample.
+ \TM{Need to improve this caption.}}
+ \label{fig:cmatrix_rest}
 \end{figure}

### 2016-05-03 Final_Project/cosmosis_code/astr_513_SNIA.py
Fixed a bug in the cosmosis code and added a routine that cuts out the burn-in. This is done by using a max-likelihood sampler before running mcmc. Make sure to change paths and switch from test sampler to emcee before running.

@@ -16,8 +16,8 @@ likes = section_names.likelihoods
 """
 adot/a \propto E_z
 """
-def E_z(z,om,ode,ok=0):
-    return np.sqrt(om*(1.0+z)**3+ok*(1.0+z)**2+ode)
+def E_z(z,om,ode,w,ok=0):
+    return np.sqrt(om*(1.0+z)**3+ok*(1.0+z)**2+ode*(1.0+z)**(3.0*(w+1)))
 
 """
 The setup function runs only once at the start
@@ -45,13 +45,13 @@ def execute(block,config):
     #Read in the cosmology
     h = block[cosmo,"h0"]
     om = block[cosmo,"omega_m"]
+    w = block[cosmo,"w"] #DE equation of state
     ode = 1.0 - om #Omega_dark_energy
 
     #Create a model for mu(z)
-    #TODO: decide how to implement this
     dc = np.ones_like(z)
     for i in range(len(z)):
-        dc[i] = 3000.0/h * integrate.quad(E_z,0,z[i],args=(om,ode))[0]
+        dc[i] = 3000.0/h * integrate.quad(E_z,0,z[i],args=(om,ode,w))[0]
     dl = (1+z)*dc
     mu_model = 5*np.log10(dl)+25
 

### 2016-05-03 Final_Project/cosmosis_code/run_script.py

@@ -14,14 +14,17 @@ if do_analysis:
     
     path = "my_tests/astr513/"
     cosmology = path+"final_cosmology.ini"
+    maxlikepipeline = path+"final_maxlike_pipeline.ini"
     pipeline = path+"final_pipeline.ini"
-    cosmosis = "cosmosis"
-    #cosmosis = "mpirun -n 4 cosmosis --mpi"
+    singlecore_cosmosis = "cosmosis"
+    mpi_cosmosis = "mpirun -n 4 cosmosis --mpi"
     
     command = "datapath=%s covpath=%s cosmology=%s %s %s"\
-              %(datapath,covpath,cosmology,cosmosis,pipeline)
-
-    os.system(command)
+              %(datapath,covpath,cosmology,singlecore_cosmosis,maxlikepipeline)
+    os.system(command) #Maxlike
+    command = "datapath=%s covpath=%s cosmology=%s %s %s"\
+              %(datapath,covpath,cosmology,mpi_cosmosis,pipeline)
+    os.system(command) #MCMC
 
 """
 Now that the code is run, we can do post processing below.

### 2016-05-03 Final_Project/ms.tex
Fixed a few typos and updated the distance modulus equation

@@
 \label{sec:SNIa}
-\EW{Should I cite the Nobel Prize information document where I found out about 
-the stuff in the last few sentences? I'm not entirely sure what that would be considered...}
-\TM{No, because it is common knowledge. I think though that since this is meant to emulate
-  a research paper that we don't need to point out that it was what the NP was awarded for.}
 A type \rom{1}a supernova (\sna{}) is a supernova with very little hydrogen or
@@
 rate of the Universe in 1998.
-%It was for this work that Saul Perlmutter, Brian P. Schmidt, 
-%and Adam G. Reiss won the Nobel Prize in physics in 2011.
-\TM{Changed grammer in this section, since ``we'' did not actually do these things.
-  Review and collapse bubbles.}
 
@@
 utilized the redshifts, distance moduli, and covariance matrix of the 
-final sample of 310 {{\snea} or 335 light curves.
-\TM{There is some kind of undefined reference error in the above line. 
-  It doesn't alter the pdf document, but can we figure out what is going on?}
+final sample of 310 {\snea} or 335 light curves.
 
@@
 The correlation matrices for both of our data sets can be seen below in
-figures \autoref{fig:cmatrix_betoule} and \autoref{cmatrix_rest}.
+figures \autoref{fig:cmatrix_betoule} and \autoref{fig:cmatrix_rest}.
 
@@
 \LC{Need more here but I wrote most of the relevant eqns}
-\EW{Do we want/need citations for the equations here?}
 The main objective of \citet{betoule2014} and \citet{rest2014} is 
@@
 \label{eq:mu_dL}
-\mu=5 \text{log}(d_L)+25. 
+\mu=5 \log\left(\frac{d_L}{1 \mathrm{Mpc}}\right)+25. 
 \end{equation}
@@
 \end{equation}
-Three (?) cosmologies are of interest to us: 
+Two cosmologies are of interest to us: 
 \LC{Not sure which ones we're interested in. $\Lambda CDM$, $\Omega_M=1$, ?}

### 2016-05-04 Final_Project/ms.tex
Wrote some things for Intro section, added a bit to cov subsection

@@
 \maketitle
+
+\section{Introduction}
+\label{sec:Intro}
+Type \rom{1}a supernovae (\snea{}) have been instrumental in the discovery of the accelerating expansion of the universe.
+Presently, {\sn} surveys continue to be pivotal in constraining cosmological parameters.
+In this paper, we present our analysis of comparing and combining {\sna} inference to those from the Cosmic Microwave Background (CMB) and the Baryonic Acoustic Oscillations (BAO).
+The constraints from these two probes are taken from \citet{planck2013}.
+This paper is outlined as follows.
+In Section 2, the process and importance of {\snea} are summarized.
+A short description of the Planck mission is given in Section 3.
+Section 4 describe the two {\sna} surveys whose data we use for our analysis.
+The significance of using distance modulus data to constrain cosmological parameters is illustrated in Section 5.
+The CMB and BAO priors taken from \citet{planck2013} are also specified in Section 5.
+Our results and analysis on the benefit of omitting or including Bayesian priors from other cosmological probes are in Section 6 and Section 7.
 
@@
 The publicly released covariance matrix from \citet{rest2014} 
-is comprised of systematic errors for the final sample. These systematics, 
-defined in \citet{scolnic2014}, include flux calibration errors, 
-selection effects, light curve fitting, host galaxy relations, Milky 
-Way extinction corrections, and coherent velocity flows. Calculating 
-the covariance matrix involves varying a systematic parameter and determining 
-the difference between the original distance and the new distance. 
-The covariance between {\sn} $i$ and {\sn} $j$ is 
+is comprised of systematic and statistical errors for the final sample.
+The total covariance matrix is defined as $\mathbf{C}= \mathbf{D}_{\textnormal{stats}}+\mathbf{C}_{\textnormal{sys}}$.
+The statistical matrix has only diagonal elements, which take into account errors in light curve fit parameters and intrinsic scatter due to luminosity and color variation.
+These statistical errors are added in quadratures.
+The systematics, defined in \citet{scolnic2014}, include flux calibration errors, selection effects, light curve fitting, host galaxy relations, Milky Way extinction corrections, and coherent velocity flows.
+Calculating the systematic matrix involves varying a systematic parameter and determining the difference between the original distance and the new distance.
+The systematic covariance between {\sn} $i$ and {\sn} $j$ is
 \begin{equation}
 \label{eq:RestCovMatrix}
-C_{ij}=\displaystyle \sum_{k = 1}^K \left(\frac{\partial \mu_i}{\partial S_k}\right)\left(\frac{\partial \mu_j}{\partial S_k}\right)\left(\sigma_{S_k}\right)^2,
+C_{ij,\textnormal{sys}}=\displaystyle \sum_{k = 1}^K \left(\frac{\partial \mu_i}{\partial S_k}\right)\left(\frac{\partial \mu_j}{\partial S_k}\right)\left(\sigma_{S_k}\right)^2,
 \end{equation}
@@
 If the variables one has measured are independent, then the correlation between
-differen variables should be $\approx0$.
+different variables should be $\approx0$.
 
@@
 \label{sec:model}
-\LC{Need more here but I wrote most of the relevant eqns}
 The main objective of \citet{betoule2014} and \citet{rest2014} is 
 to place constraints on cosmological parameters $\Omega_M$, $\Omega_\Lambda$, 
-and $w$. In order to so, we must examine the relationship between distance 
+$\Omega_k$, and $w$. In order to so, we must examine the relationship between distance 
 modulus and cosmology. The distance modulus $\mu$ is determined from 
@@
 \end{equation}
-Two cosmologies are of interest to us: 
-\LC{Not sure which ones we're interested in. $\Lambda CDM$, $\Omega_M=1$, ?}
-\TM{I think there are only two, one with $\Omega_k=0$ and one with it
-  being unconstrained.}
+Two cosmologies are of interest to us: a flat universe where $\Omega_k=0$ and a universe where $\Omega_k$ is unconstrained.
 

### 2016-05-04 Final_Project/cosmosis_code/quick_plot.py
Added a quick script that plots the data and a model.

@@ -1,0 +1,44 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy import integrate
+
+#MODEL:
+om = 0.3
+h = 0.69
+w = -1.
+ode = 1. - om
+
+def E_z(z,om,ode,w,ok=0):
+    return np.sqrt(om*(1.0+z)**3.0+ok*(1.0+z)**2.0+ode*(1.0+z)**(3.0*(w+1.0)))
+
+betoule = False
+rest = True
+
+if betoule:
+    datapath = "../Betoule2014Data/tablef1.dat"
+    covpath = "../Betoule2014Data/tablef2.dat"
+elif rest:
+    datapath = "../Rest2014Data/redshift_distanceMod.dat"
+    covpath = "../Rest2014Data/covarianceMatrix.dat"
+
+data = np.genfromtxt(datapath)
+cov = np.genfromtxt(covpath)
+errs = np.sqrt(np.diag(cov)) #errorbars
+icov = np.linalg.inv(cov)
+z,mu = data.T
+inds = np.argsort(z)
+z = z[inds]
+mu = mu[inds]
+
+#Create a model for mu(z)
+dc = np.ones_like(z)
+for i in range(len(z)):
+    dc[i] = 3000.0/h * integrate.quad(E_z,0,z[i],args=(om,ode,w))[0]
+dl = (1+z)*dc
+mu_model = 5*np.log10(dl)+25
+    
+plt.errorbar(z,mu,yerr=errs,ls='',marker='.',alpha=0.5)
+plt.plot(z,mu_model)
+#plt.xscale('log')
+#plt.yscale('log')
+plt.show()

### 2016-05-05 Final_Project/ms.tex
Modified intro and distance mod sections a bit. Added paragraph to results section.

@@
 Presently, {\sn} surveys continue to be pivotal in constraining cosmological parameters.
-In this paper, we present our analysis of comparing and combining {\sna} inference to those from the Cosmic Microwave Background (CMB) and the Baryonic Acoustic Oscillations (BAO).
-The constraints from these two probes are taken from \citet{planck2013}.
+In this paper, we present our analysis of comparing and combining {\sna} inference of cosmological parameters to those from other cosmological probes.
+In particular, we utilize the constraints from the Cosmic Microwave Background (CMB) and the Baryonic Acoustic Oscillations (BAO) that are gathered in \citet{planck2013}.
+We look at two cosmologies in our analysis: the $\Lambda \textnormal{CDM}$ universe and the wCDM universe.
 This paper is outlined as follows.
@@
 \label{sec:SNIa}
-A type \rom{1}a supernova (\sna{}) is a supernova with very little hydrogen or
+A type \rom{1}a supernova is a supernova with very little hydrogen or
 helium absorption. It is believed that the progenitor for these special
@@
 In 2009 the European Space Agency launched the \Planck satellite in order
-to measure the Cosmis Microwave Background (CMB) with unparalleled precision.
+to measure the Cosmis Microwave Background with unparalleled precision.
 By measuring the small anisotropies in the temperature of the CMB across
@@
     \text{sinh}(x),& \Omega_k > 0
-\end{cases}
+\end{cases}.
 \]
-and the function $H(z)$ is 
+The function $H(z)$ is expressed as 
 \begin{equation}
 \label{eq:hubbleFunction}
-H(z)= H_0\displaystyle\left[\Omega_M(1+z)^3+\Omega_{\Lambda}(1+z)^{3(1+w)}+\Omega_k(1+z)^2\right]^{1/2}.
-\end{equation}
-Two cosmologies are of interest to us: a flat universe where $\Omega_k=0$ and a universe where $\Omega_k$ is unconstrained.
+H(z)= H_0\displaystyle\left[\Omega_M(1+z)^3+\Omega_{\Lambda}(1+z)^{3(1+w)}+\Omega_k(1+z)^2\right]^{1/2},
+\end{equation} where the Hubble constant $H_0$ can be written in terms of a dimensionless Hubble parameter $h$ by $H_0=100h \mathrm{kms^{-1}Mpc^{-1}}$.
+
+Two cosmologies are of interest to us: the $\Lambda \textnormal{CDM}$ universe and the wCDM universe.
+In the $\Lambda \textnormal{CDM}$ universe, $w=-1$ and constraints are placed on $\Omega_M$ and $h$.
+The wCDM universe is a one-parameter extension of $\Lambda \textnormal{CDM}$ in which $w$ is an arbitrary parameter for a constant dark energy equation of state.
+In wCDM, constraints are placed on $\Omega_M$, $w$, and $h$.
+In both cosmologies, the universe is spatially flat ($\Omega_k = 0$) and thus $\Omega_{\Lambda}=1-\Omega_M$.
 
@@
   We should have one corner plot without priors and one with priors.}
+\LC{Not sure where to put why we couldn't do the full analysis of Rest's data. I'm putting it here for now.}
+A complete analysis using the data from \citet{rest2014} is not done in our paper since there are several issues with the publicly released data.
+It is not apparent which uncertainties belonged to the statistical covariance matrix since only the systematic covariance matrix is clearly presented.
+Moreover, intrinsic dispersion of the data and photometric errors of the distances depend on nuisance parameters in addition to the covariances between the fit parameters.
+Fine-tuning nuisance parameters places a significant time restraint on our work and is beyond the scope of this paper.
+For these reasons, we only present the results for \citet{rest2014} that do not include the priors from \citet{planck2013}.
 

### 2016-05-05 Final_Project/ms.tex
Wrote the likelihood section.

@@
 \label{sec:Intro}
-Type \rom{1}a supernovae (\snea{}) have been instrumental in the discovery of the accelerating expansion of the universe.
-Presently, {\sn} surveys continue to be pivotal in constraining cosmological parameters.
-In this paper, we present our analysis of comparing and combining {\sna} inference of cosmological parameters to those from other cosmological probes.
-In particular, we utilize the constraints from the Cosmic Microwave Background (CMB) and the Baryonic Acoustic Oscillations (BAO) that are gathered in \citet{planck2013}.
-We look at two cosmologies in our analysis: the $\Lambda \textnormal{CDM}$ universe and the wCDM universe.
-This paper is outlined as follows.
-In Section 2, the process and importance of {\snea} are summarized.
-A short description of the Planck mission is given in Section 3.
-Section 4 describe the two {\sna} surveys whose data we use for our analysis.
-The significance of using distance modulus data to constrain cosmological parameters is illustrated in Section 5.
-The CMB and BAO priors taken from \citet{planck2013} are also specified in Section 5.
-Our results and analysis on the benefit of omitting or including Bayesian priors from other cosmological probes are in Section 6 and Section 7.
+Type \rom{1}a supernovae (\snea{}) have been instrumental in the discovery 
+of the accelerating expansion of the universe. Presently, {\sn} surveys continue 
+to be pivotal in constraining cosmological parameters. In this paper, we present 
+our analysis of comparing and combining {\sna} inference of cosmological parameters 
+to those from other cosmological probes. In particular, we utilize the constraints 
+from the Cosmic Microwave Background (CMB) and the Baryonic Acoustic Oscillations 
+(BAO) that are gathered in \citet{planck2013}. We look at two cosmologies in our 
+analysis: the $\Lambda \textnormal{CDM}$ universe and the wCDM universe. This 
+paper is outlined as follows. In Section 2, the process and importance of {\snea} 
+are summarized. A short description of the Planck mission is given in Section 3. 
+Section 4 describe the two {\sna} surveys whose data we use for our analysis. The 
+significance of using distance modulus data to constrain cosmological parameters 
+is illustrated in Section 5. The CMB and BAO priors taken from \citet{planck2013} 
+are also specified in Section 5. Our results and analysis on the benefit of omitting 
+or including Bayesian priors from other cosmological probes 
+are in Section 6 and Section 7.
 
@@
 is comprised of systematic and statistical errors for the final sample. 
-The total covariance matrix is defined as $\mathbf{C}= \mathbf{D}_{\textnormal{stats}}+\mathbf{C}_{\textnormal{sys}}$.
-The statistical matrix has only diagonal elements, which take into account errors in light curve fit parameters and intrinsic scatter due to luminosity and color variation.
-These statistical errors are added in quadratures.
-The systematics, defined in \citet{scolnic2014}, include flux calibration errors, selection effects, light curve fitting, host galaxy relations, Milky Way extinction corrections, and coherent velocity flows.
-Calculating the systematic matrix involves varying a systematic parameter and determining the difference between the original distance and the new distance.
-The systematic covariance between {\sn} $i$ and {\sn} $j$ is
+The total covariance matrix is defined as 
+$\mathbf{C}= \mathbf{D}_{\textnormal{stats}}+\mathbf{C}_{\textnormal{sys}}$. The 
+statistical matrix has only diagonal elements, which take into account errors 
+in light curve fit parameters and intrinsic scatter due to luminosity and 
+color variation. These statistical errors are added in quadratures. The 
+systematics, defined in \citet{scolnic2014}, include flux calibration errors, 
+selection effects, light curve fitting, host galaxy relations, Milky Way 
+extinction corrections, and coherent velocity flows. Calculating the systematic 
+matrix involves varying a systematic parameter and determining the difference 
+between the original distance and the new distance. The systematic covariance 
+between {\sn} $i$ and {\sn} $j$ is 
 \begin{equation}
@@
 H(z)= H_0\displaystyle\left[\Omega_M(1+z)^3+\Omega_{\Lambda}(1+z)^{3(1+w)}+\Omega_k(1+z)^2\right]^{1/2},
-\end{equation} where the Hubble constant $H_0$ can be written in terms of a dimensionless Hubble parameter $h$ by $H_0=100h \mathrm{kms^{-1}Mpc^{-1}}$.
+\end{equation} where the Hubble constant $H_0$ can be written in terms of a 
+dimensionless Hubble parameter $h$ by $H_0=100h \mathrm{km\ s^{-1}Mpc^{-1}}$.
 
-Two cosmologies are of interest to us: the $\Lambda \textnormal{CDM}$ universe and the wCDM universe.
-In the $\Lambda \textnormal{CDM}$ universe, $w=-1$ and constraints are placed on $\Omega_M$ and $h$.
-The wCDM universe is a one-parameter extension of $\Lambda \textnormal{CDM}$ in which $w$ is an arbitrary parameter for a constant dark energy equation of state.
-In wCDM, constraints are placed on $\Omega_M$, $w$, and $h$.
-In both cosmologies, the universe is spatially flat ($\Omega_k = 0$) and thus $\Omega_{\Lambda}=1-\Omega_M$.
+Two cosmologies are of interest to us: the $\Lambda \textnormal{CDM}$ 
+universe and the wCDM universe. In the $\Lambda \textnormal{CDM}$ 
+universe, $w=-1$ and constraints are placed on $\Omega_M$ and $h$. 
+The wCDM universe is a one-parameter extension of $\Lambda \textnormal{CDM}$ 
+in which $w$ is an arbitrary parameter for a constant dark energy 
+equation of state. In wCDM, constraints are placed on $\Omega_M$, $w$, 
+and $h$. In both cosmologies, the universe is spatially flat 
+($\Omega_k = 0$) and thus $\Omega_{\Lambda}=1-\Omega_M$.
 
@@
 \label{sec:likelihood}
-\TM{This has our likelihood.}
+Using the above model for the distance modulus allows us to write a
+probability that a model can generate our data:
+%
+\begin{equation}
+  \label{eq:likelihood}
+  P(\mu(z)|\mathrm{cos}) = \frac{1}{\sqrt{2\pi|\mathbf{C}|}}\exp
+  \left[-\frac{1}{2}\sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i\mathbf{C}^{-1}_{i,j}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_j\right]
+\end{equation}
+%
+where $\mathbf{C}$ is the full covariance matrix between all of the $N_{samples}$ 
+in the data, and $\mathrm{cos}$ represents the cosmology that we are investigating
+(e.g. flat $\Lambda$CDM, $wCDM$, etc.). 
+This is not the full picture though, since we must employ
+Bayes' Theorum in order to include any priors on our choice of 
+parameters and recover the posterior distribution function of our
+parameters given our data
+%
+\begin{equation}
+  \label{eq:posterior}
+  P(\mathrm{cos}|\mu(z)) = P_{prior}(\mathrm{cos}) P(\mu(z)|\mathrm{cos}).
+\end{equation}
+%
+For this work we investigate a combination of priors. \Cosmosis is implemented
+in a way that requires the user to provide bounds on the parameter space, which we
+detail for each parameter in Table~XXX
+\TM{Make this table. Just look at the cosmology.ini files.}
+Within those bounds, we investigated using a flat prior $P_{prior}=1$ and
+\Planck priors.
 
@@
 \label{sec:planck_priors}
+\TM{The only thing we really require is a corner plot of {\it just} the priors.}
 
@@
   We should have one corner plot without priors and one with priors.}
-\LC{Not sure where to put why we couldn't do the full analysis of Rest's data. I'm putting it here for now.}
-A complete analysis using the data from \citet{rest2014} is not done in our paper since there are several issues with the publicly released data.
-It is not apparent which uncertainties belonged to the statistical covariance matrix since only the systematic covariance matrix is clearly presented.
-Moreover, intrinsic dispersion of the data and photometric errors of the distances depend on nuisance parameters in addition to the covariances between the fit parameters.
-Fine-tuning nuisance parameters places a significant time restraint on our work and is beyond the scope of this paper.
-For these reasons, we only present the results for \citet{rest2014} that do not include the priors from \citet{planck2013}.
+\LC{Not sure where to put why we couldn't do the full analysis of Rest's data. 
+  I'm putting it here for now.}
+A complete analysis using the data from \citet{rest2014} is not done 
+in our paper since there are several issues with the publicly released 
+data. It is not apparent which uncertainties belonged to the statistical 
+covariance matrix since only the systematic covariance matrix is clearly 
+presented. Moreover, intrinsic dispersion of the data and photometric errors 
+of the distances depend on nuisance parameters in addition to the covariances 
+between the fit parameters. Fine-tuning nuisance parameters places a significant 
+time restraint on our work and is beyond the scope of this paper. For these reasons, 
+we only present the results for \citet{rest2014} that do not include the priors 
+from \citet{planck2013}.
 

### 2016-05-05 Final_Project/ms.tex
Wrote the likelihood section.

@@
 %
-\begin{equation}
+\begin{eqnarray}
   \label{eq:likelihood}
-  P(\mu(z)|\mathrm{cos}) = \frac{1}{\sqrt{2\pi|\mathbf{C}|}}\exp
-  \left[-\frac{1}{2}\sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i\mathbf{C}^{-1}_{i,j}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_j\right]
-\end{equation}
+  P(\mu(z)|\mathrm{cosmo}) = \frac{1}{\sqrt{2\pi|\mathbf{C}|}}\exp
+  \left[-\frac{1}{2}& \\ \sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i 
+    \mathbf{C}^{-1}_{i,j}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_j\right]
+\end{eqnarray}
 %
 where $\mathbf{C}$ is the full covariance matrix between all of the $N_{samples}$ 
-in the data, and $\mathrm{cos}$ represents the cosmology that we are investigating
+in the data, and $\mathrm{cosmo}$ represents the cosmology that we are investigating
 (e.g. flat $\Lambda$CDM, $wCDM$, etc.). 
@@
   \label{eq:posterior}
-  P(\mathrm{cos}|\mu(z)) = P_{prior}(\mathrm{cos}) P(\mu(z)|\mathrm{cos}).
+  P(\mathrm{cosmo}|\mu(z)) = P_{prior}(\mathrm{cosmo}) P(\mu(z)|\mathrm{cosmo}).
 \end{equation}

### 2016-05-05 Final_Project/ms.tex
Added some whitespace to make equations easier to read.

@@
 \citep[see][Appendix E.1]{betoule2014}
+%
 \begin{equation}
@@
 \end{equation}
+%
 where they have defined 
@@
 between {\sn} $i$ and {\sn} $j$ is 
+%
 \begin{equation}
@@
 \end{equation}
+%
 where the summation is over $K$ systematics, $\sigma_{S_k}$ is each 
@@
 terms to account for uncertainty in cosmological redshift by the equation
+%
 \begin{equation}
@@
 \end{equation}
+%
 where $\mathbf{A}$ is the matrix of coefficients in the distance modulus, 
@@
 The correlation matrix is calculated from the covariance matrix according to
+%
 \begin{equation}
@@
 \end{equation}
+%
 Note that the correlation of a variable with itself is 1 by definition.
@@
 the luminosity distance $d_L$ by 
+%
 \begin{equation}
@@
 \end{equation}
+%
 The luminosity distance depends on the cosmology by 
+%
 \begin{equation}
@@
 \end{equation}
+%
 where $\Omega_k \equiv 1-\Omega_M-\Omega_{\Lambda}$. The function 
@@
 The function $H(z)$ is expressed as 
+%
 \begin{equation}
@@
   H(z)= H_0\displaystyle\left[\Omega_M(1+z)^3+\Omega_{\Lambda}(1+z)^{3(1+w)}+\Omega_k(1+z)^2\right]^{1/2},
-\end{equation} where the Hubble constant $H_0$ can be written in terms of a 
+\end{equation}
+%
+where the Hubble constant $H_0$ can be written in terms of a 
 dimensionless Hubble parameter $h$ by $H_0=100h \mathrm{km\ s^{-1}Mpc^{-1}}$.
@@
   P(\mu(z)|\mathrm{cosmo}) = \frac{1}{\sqrt{2\pi|\mathbf{C}|}}\exp
-  \left[-\frac{1}{2}& \\ \sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i 
+  \left[-\frac{1}{2} \sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i 
     \mathbf{C}^{-1}_{i,j}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_j\right]
@@
 in a way that requires the user to provide bounds on the parameter space, which we
-detail for each parameter in Table~XXX
-\TM{Make this table. Just look at the cosmology.ini files.}
+detail for each parameter in \autoref{tab:priors}
 Within those bounds, we investigated using a flat prior $P_{prior}=1$ and
 \Planck priors.
+
+\begin{table}
+  \centering
+  \caption{Bounds on cosmological parameters 
+    entering \autoref{eq:posterior}. Our sampler requires
+    bounds in order to both populate the parameter space with
+    parallelized samplers and to calculate an appropriate width
+    for the step size.}
+  \label{tab:priors}
+  \begin{tabular}{lll}
+    Parameter & Description & Bounds \\ \hline
+    $h$ & Hubble constant$/100$ & $[0,2]$ \\
+    $\Omega_M$ & Matter fraction at $z=0$ & $[0,1]$ \\
+    $w$ & Dark energy E.O.S. & $[-2,0]$\\
+  \end{tabular}
+\end{table}
 

### 2016-05-05 Final_Project/ms.tex
Redcued the width of the correlation matrix figures so that they don't each take up a page.
Improved figure captions.

@@
 figures \autoref{fig:cmatrix_betoule} and \autoref{fig:cmatrix_rest}.
-
-\TM{TODO: Discuss how the correlation matrices for the two data sets are different
-  and hypothesize on what this might mean.}
-
+%
 \begin{figure}
- \includegraphics[width=\linewidth]{figures/Betoule_correlation.eps}
+  \includegraphics[width=0.8\linewidth]{figures/Betoule_correlation.eps}
   \caption{Correlation matrix of $\mu_0(z)$ from the
- \citet{betoule2014} sample.
- \TM{Need to improve this caption.}}
+    \citet{betoule2014} sample. Overall very few correlations exist between
+    samples, likely due to the averaging over redshift bins.}
   \label{fig:cmatrix_betoule}
 \end{figure}
-
+%
 \begin{figure}
- \includegraphics[width=\linewidth]{figures/Rest_correlation.eps}
+  \includegraphics[width=0.8\linewidth]{figures/Rest_correlation.eps}
   \caption{Correlation matrix of $\mu_0(z)$ from the
- \citet{rest2014} sample.
- \TM{Need to improve this caption.}}
+    \citet{rest2014} sample. Significant correlation exist between
+    different samples due to dependencies on systematic variables
+    as seen in \autoref{eq:RestCovMatrix}.}
   \label{fig:cmatrix_rest}
@@
 \begin{figure}
- \includegraphics[width=\linewidth]{figures/corner_example.eps}
+  \centering
+  \includegraphics[width=0.5\linewidth]{figures/corner_example.eps}
   \caption{Example corner plot.}

### 2016-05-05 Final_Project/ms.tex
Wrote likelihood. Fixed figure captions. Put in line breaks. Added an appendix.

@@
   \centering
-  \includegraphics[width=0.5\linewidth]{figures/corner_example.eps}
-  \caption{Example corner plot.}
+  \includegraphics[width=0.8\linewidth]{figures/corner_example.eps}
+  \caption{\TM{THIS IS A PLACE HOLDER PLOT!}
+    Posterior distributions of the cosmological parameters in the
+    analysis of the \citet{Betoule2014} data. Contours denote the
+    68\% and 95\% confidence projected areas; dashed lines in
+    the 1D histograms refer to the 68\% confidence intervals.}
   \label{fig:corner}
@@
 
+\appendix
+\section{Posterior Distribution Functions Without \Planck Priors}
+\TM{Since there are potentially a bunch of corner plots, we could jam a bunch of
+  them here as well as perhaps the discussion of what happens when we remove
+  the priros. Let me know your thoughts.}
+
 \end{document}

### 2016-05-06 Final_Project/cosmosis_code/astr_513_SNIA.py
Found possible bug in SNIA code
Think I may have found a bug in the code, so I added what I think it
should be as a comment. Take a look and tell me what you guys think.

@@ -51,7 +51,9 @@ def execute(block,config):
     #Create a model for mu(z)
     dc = np.ones_like(z)
     for i in range(len(z)):
+        ## This looks wrong to me -- ELW
         dc[i] = 3000.0/h * integrate.quad(E_z,0,z[i],args=(om,ode,w))[0]
+        #dc[i] = 3000.0/h * integrate.quad(1./E_z,0,z[i],args=(om,ode,w))[0]
     dl = (1+z)*dc
     mu_model = 5*np.log10(dl)+25
 

### 2016-05-06 Final_Project/cosmosis_code/astr_513_SNIA.py
Added 1/E(z) function

@@ -18,6 +18,8 @@ adot/a \propto E_z
 """
 def E_z(z,om,ode,w,ok=0):
     return np.sqrt(om*(1.0+z)**3+ok*(1.0+z)**2+ode*(1.0+z)**(3.0*(w+1)))
+def E_z_inv(z,om,ode,w,ok=0):
+    return 1./E_z(z,om,ode,w,ok)
 
 """
 The setup function runs only once at the start
@@ -53,7 +55,7 @@ def execute(block,config):
     for i in range(len(z)):
         ## This looks wrong to me -- ELW
         dc[i] = 3000.0/h * integrate.quad(E_z,0,z[i],args=(om,ode,w))[0]
-        #dc[i] = 3000.0/h * integrate.quad(1./E_z,0,z[i],args=(om,ode,w))[0]
+        #dc[i] = 3000.0/h * integrate.quad(E_z_inv,0,z[i],args=(om,ode,w))[0]
     dl = (1+z)*dc
     mu_model = 5*np.log10(dl)+25
 

### 2016-05-06 Final_Project/ms.tex
Changed little things, added a bit about cosmosis

@@
 \newcommand\Planck{{\it Planck}\ }
-\newcommand\Cosmosis{{\it Cosmosis}\ }
+\newcommand\CosmoSIS{{\it CosmoSIS}\ }
 
@@
 paper is outlined as follows. In Section 2, the process and importance of {\snea} 
-are summarized. A short description of the Planck mission is given in Section 3. 
+are summarized. A short description of the \Planck mission is given in Section 3. 
 Section 4 describe the two {\sna} surveys whose data we use for our analysis. The 
@@
 which we report in Table~XXX. This prior is implemented
-in the planck2015 module implemented in \Cosmosis \citep{cosmosis}.
+in the planck2015 module implemented in \CosmoSIS \citep{cosmosis}.
 
@@
 
-Two cosmologies are of interest to us: the $\Lambda \textnormal{CDM}$ 
-universe and the wCDM universe. In the $\Lambda \textnormal{CDM}$ 
+Two cosmologies are of interest to us: the  $\Lambda$CDM 
+universe and the $w$CDM universe. In the  $\Lambda$CDM 
 universe, $w=-1$ and constraints are placed on $\Omega_M$ and $h$. 
-The wCDM universe is a one-parameter extension of $\Lambda \textnormal{CDM}$ 
+The $w$CDM universe is a one-parameter extension of  $\Lambda$CDM 
 in which $w$ is an arbitrary parameter for a constant dark energy 
-equation of state. In wCDM, constraints are placed on $\Omega_M$, $w$, 
+equation of state. In $w$CDM, constraints are placed on $\Omega_M$, $w$, 
 and $h$. In both cosmologies, the universe is spatially flat 
 ($\Omega_k = 0$) and thus $\Omega_{\Lambda}=1-\Omega_M$.
+
+\subsection{CosmoSIS}
+\label{sec:CosmoSIS}
+\LC{I'll write more later if need be.}
+\CosmoSIS is a framework that connects multiple existing codes as modules in order to do cosmological parameter estimation.
+Users can utilize existing codes or use their own codes for modules.
+For our analysis, \CosmoSIS provides a convenient way to link together the Code for Anisotropies in the Microwave Background (CAMB), the \Planck likelihood code (our priors), and our {\sn} likelihood code in a pipeline.
 
@@
 in the data, and $\mathrm{cosmo}$ represents the cosmology that we are investigating
-(e.g. flat $\Lambda$CDM, $wCDM$, etc.). 
+(e.g. flat $\Lambda$CDM, $w$CDM, etc.). 
 This is not the full picture though, since we must employ
@@
 %
-For this work we investigate a combination of priors. \Cosmosis is implemented
+For this work we investigate a combination of priors. \CosmoSIS is implemented
 in a way that requires the user to provide bounds on the parameter space, which we

### 2016-05-06 Final_Project/ms.tex
planck, cosmosis

@@
 for the optical depth $\tau$. We do not use this parameter
-in our {\sna} model.
-
-\citet{planck2013} report marginalized
+in our {\sna} model.\citet{planck2013} report marginalized
 posterior distributions for all parameters,
-which we report in Table~XXX. This prior is implemented
-in the planck2015 module implemented in \CosmoSIS \citep{cosmosis}.
+which we report in Table~XXX.
 

### 2016-05-06 Final_Project/cosmosis_code/astr_513_SNIA.py
Vectorized a few things in the module

@@ -20,6 +20,9 @@ def E_z(z,om,ode,w,ok=0):
     return np.sqrt(om*(1.0+z)**3+ok*(1.0+z)**2+ode*(1.0+z)**(3.0*(w+1)))
 def E_z_inv(z,om,ode,w,ok=0):
     return 1./E_z(z,om,ode,w,ok)
+def dc(h,z,om,ode,w,ok=0):
+    return 3000.0/h * integrate.quad(E_z_inv,0,z,args=(om,ode,w,ok))[0]
+dc = np.vectorize(dc)
 
 """
 The setup function runs only once at the start
@@ -51,19 +54,16 @@ def execute(block,config):
     ode = 1.0 - om #Omega_dark_energy
 
     #Create a model for mu(z)
-    dc = np.ones_like(z)
-    for i in range(len(z)):
+    #dc = np.ones_like(z)
+    #for i in range(len(z)):
         ## This looks wrong to me -- ELW
-        dc[i] = 3000.0/h * integrate.quad(E_z,0,z[i],args=(om,ode,w))[0]
+        #dc[i] = 3000.0/h * integrate.quad(E_z,0,z[i],args=(om,ode,w))[0]
         #dc[i] = 3000.0/h * integrate.quad(E_z_inv,0,z[i],args=(om,ode,w))[0]
-    dl = (1+z)*dc
+    dl = (1+z)*dc(z)
     mu_model = 5*np.log10(dl)+25
 
     #Calculate the likelihood
-    LL = 0.0
-    for i in range(len(cov)):
-        for j in range(len(cov[i])):
-            LL += -0.5 * (mu[i]-mu_model[i])*icov[i,j]*(mu[j]-mu_model[j])
+    LL = -0.5 np.dot(mu-mu_model,np.dot(icov,mu-mu_model))
     block[likes,"SNIA_LIKE"]=LL
 
     return 0 #Signal that it finished fine

### 2016-05-06 Final_Project/cosmosis_code/quick_plot.py
Added corner files.

@@ -4,12 +4,12 @@ from scipy import integrate
 
 #MODEL:
 om = 0.3
-h = 0.69
-w = -1.
+h = 0.5
+w = 0.0
 ode = 1. - om
 
 def E_z(z,om,ode,w,ok=0):
-    return np.sqrt(om*(1.0+z)**3.0+ok*(1.0+z)**2.0+ode*(1.0+z)**(3.0*(w+1.0)))
+    return 1./np.sqrt(om*(1.0+z)**3.0+ok*(1.0+z)**2.0+ode*(1.0+z)**(3.0*(w+1.0)))
 
 betoule = False
 rest = True
@@ -29,6 +29,8 @@ z,mu = data.T
 inds = np.argsort(z)
 z = z[inds]
 mu = mu[inds]
+plt.imshow(cov)
+plt.show()
 
 #Create a model for mu(z)
 dc = np.ones_like(z)
@@ -41,4 +43,5 @@ plt.errorbar(z,mu,yerr=errs,ls='',marker='.',alpha=0.5)
 plt.plot(z,mu_model)
 #plt.xscale('log')
 #plt.yscale('log')
+#plt.ylim([3,5])
 plt.show()

### 2016-05-06 Final_Project/ms.tex
Added filler tables to the tex file for fit results

@@
   \caption{Correlation matrix of $\mu_0(z)$ from the
-    \citet{rest2014} sample. Significant correlation exist between
+    \citet{rest2014} sample. Significant correlation exists between
     different samples due to dependencies on systematic variables
@@
 
+\EW{Filler table for the lambda CDM fits}
+\begin{table}[h]
+   \centering
+    \begin{tabular}{l|cc}
+    ~                          & $\Omega_m$ & $h$ \\ \hline
+    Betoule et al.             & 0      & 0 \\
+    Betoule et al. (no priors) & 0      & 0 \\
+    Rest et al.                & 0      & 0 \\
+    Rest et al. (no priors)    & 0      & 0 \\
+    \end{tabular}
+    \caption {\label{tab:lcdm}$\Lambda$CDM best fit values}
+\end{table}
+
+\EW{Just a filler for the table for the wCDM fits}
+\begin{table}[h]
+   \centering
+    \begin{tabular}{l|ccc}
+    ~                          & $\Omega_m$ & $h$ & $w$ \\ \hline
+    Betoule et al.             & 0                      & 0     & 0     \\
+    Betoule et al. (no priors) & 0                      & 0     & 0     \\
+    Rest et al.                & 0                      & 0     & 0     \\
+    Rest et al. (no priors)    & 0                      & 0     & 0     \\
+    \end{tabular}
+    \caption {\label{tab:wcdm}wCDM best fit values}
+\end{table}
+
 \begin{figure}

### 2016-05-07 Final_Project/ms.tex
Renamed some files and put in figures into the paper. Also wrote more.

@@
 
-\subsection{CosmoSIS}
-\label{sec:CosmoSIS}
-\LC{I'll write more later if need be.}
-\CosmoSIS is a framework that connects multiple existing codes as modules in order to do cosmological parameter estimation.
-Users can utilize existing codes or use their own codes for modules.
-For our analysis, \CosmoSIS provides a convenient way to link together the Code for Anisotropies in the Microwave Background (CAMB), the \Planck likelihood code (our priors), and our {\sn} likelihood code in a pipeline.
-
 \subsection{Likelihood}
@@
 %
-\begin{eqnarray}
+%\begin{equation}
+\begin{align}
   \label{eq:likelihood}
-  P(\mu(z)|\mathrm{cosmo}) = \frac{1}{\sqrt{2\pi|\mathbf{C}|}}\exp
-  \left[-\frac{1}{2} \sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i 
+  P(\mu(z)|\mathrm{cosmo}) &= \frac{1}{\sqrt{2\pi|\mathbf{C}|}}\nonumber \\
+  &\qquad{}\exp\left[-\frac{1}{2} \sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i 
     \mathbf{C}^{-1}_{i,j}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_j\right]
-\end{eqnarray}
+\end{align}
+%\end{equation}
 %
@@
 \label{sec:planck_priors}
-\TM{The only thing we really require is a corner plot of {\it just} the priors.}
+Our priors are given by the analysis of the \citet{planck2013} data set.
+However, instead of simply using the posterior mean and variance of
+the parameters from that analysis as priors in our project, we
+employed the full \Planck likelihood as a prior. That is,
+we first run the Code for Anisotropies in the Microwave Background (CAMB)
+in order to obtain the temperature power spectrum $\mathcal{D}_l^{TT}$
+and the polarization power spectrum $\mathcal{C}_l^{EE}$
+and feed it into the \Planck likelihood code\footnote{More information
+on the \Planck likelihood can be found here: 
+\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/CMB_spectrum_\%26_Likelihood_Code}.}.
+The resulting likelihood
+is our prior, and it captures any degeneracies in parameter space
+observed by the \Planck analysis, as seen in figure 3.
+%\autoref{fig:planck_corner}
+\TM{The autoref isn't working here. If someone can figure it out put it back in.}
+
+\begin{figure}
+  \label{fig:planck_corner}
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/corner_example.eps}
+  \caption{\TM{THIS IS A PLACE HOLDER PLOT!}
+    Posterior distributions of the cosmological parameter
+    analysis using just the \Planck likelihood. Contours denote the
+    68\% and 95\% confidence projected areas; dashed lines in
+    the 1D histograms refer to the 68\% confidence intervals.}
+\end{figure}
+
+\subsection{CosmoSIS}
+\label{sec:CosmoSIS}
+\LC{I'll write more later if need be.}
+\CosmoSIS is a framework that connects multiple existing codes as 
+modules in order to do cosmological parameter estimation. Users 
+can utilize existing codes or use their own codes for modules. For 
+our analysis, \CosmoSIS provides a convenient way to link together 
+CAMB, the \Planck likelihood code (our priors), and our {\sn} likelihood 
+code in a single pipeline. This design choice minimized the significant
+time that would inevitably have been spent on connecting these 
+independently-made programs into one, allowing us to focus on 
+doing science rather than software development.
 
@@
   We should have one corner plot without priors and one with priors.}
-\LC{Not sure where to put why we couldn't do the full analysis of Rest's data. 
-  I'm putting it here for now.}
-A complete analysis using the data from \citet{rest2014} is not done 
-in our paper since there are several issues with the publicly released 
-data. It is not apparent which uncertainties belonged to the statistical 
-covariance matrix since only the systematic covariance matrix is clearly 
-presented. Moreover, intrinsic dispersion of the data and photometric errors 
-of the distances depend on nuisance parameters in addition to the covariances 
-between the fit parameters. Fine-tuning nuisance parameters places a significant 
-time restraint on our work and is beyond the scope of this paper. For these reasons, 
-we only present the results for \citet{rest2014} that do not include the priors 
-from \citet{planck2013}.
+%\LC{Not sure where to put why we couldn't do the full analysis of Rest's data. 
+%  I'm putting it here for now.}
+%A complete analysis using the data from \citet{rest2014} is not done 
+%in our paper since there are several issues with the publicly released 
+%data. It is not apparent which uncertainties belonged to the statistical 
+%covariance matrix since only the systematic covariance matrix is clearly 
+%presented. Moreover, intrinsic dispersion of the data and photometric errors 
+%of the distances depend on nuisance parameters in addition to the covariances 
+%between the fit parameters. Fine-tuning nuisance parameters places a significant 
+%time restraint on our work and is beyond the scope of this paper. For these reasons, 
+%we only present the results for \citet{rest2014} that do not include the priors 
+%from \citet{planck2013}.
 
@@
     \end{tabular}
-    \caption {\label{tab:lcdm}$\Lambda$CDM best fit values}
+    \caption {\label{tab:lcdm}Posterior distirbutions of parameters in the
+      $\Lambda$CDM model. uncertainties denote symmetrized 68\%
+      confidence intervals maginalized over other parameters.}
 \end{table}
@@
 \begin{table}[h]
+  \label{tab:wcdm}
   \centering
@@
   \end{tabular}
-    \caption {\label{tab:wcdm}wCDM best fit values}
+  \caption {Posterior distirbutions of parameters in the
+    wCDM model. Uncertainties denote symmetrized 68\%
+    confidence intervals maginalized over other parameters.}
 \end{table}
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/corner_example.eps}
-  \caption{\TM{THIS IS A PLACE HOLDER PLOT!}
-    Posterior distributions of the cosmological parameters in the
-    analysis of the \citet{Betoule2014} data. Contours denote the
+  \includegraphics[width=0.8\linewidth]{figures/betoule_wcdm_noprior_corner.eps}
+  \caption{Posterior distributions of the cosmological parameters in the
+    analysis of the \citet{Betoule2014} data without including the
+    \Planck priors. Contours denote the
     68\% and 95\% confidence projected areas; dashed lines in
     the 1D histograms refer to the 68\% confidence intervals.}
-  \label{fig:corner}
+  \label{fig:betoule_wcdm_noprior_corner}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/betoule_wcdm_withprior_corner.eps}
+  \caption{Posterior distributions of the cosmological parameters in the
+    analysis of the \citet{Betoule2014} data including the
+    \Planck priors. Contours denote the
+    68\% and 95\% confidence projected areas; dashed lines in
+    the 1D histograms refer to the 68\% confidence intervals.}
+  \label{fig:betoule_wcdm_withprior_corner}
 \end{figure}

### 2016-05-07 Final_Project/ms.tex
Updated captions.

@@
 %
-%\begin{equation}
 \begin{align}
@@
 \end{align}
-%\end{equation}
 %
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/corner_example.eps}
+  \includegraphics[width=0.6\linewidth]{figures/corner_example.eps}
   \caption{\TM{THIS IS A PLACE HOLDER PLOT!}
@@
 \label{sec:CosmoSIS}
-\LC{I'll write more later if need be.}
 \CosmoSIS is a framework that connects multiple existing codes as 
@@
     Rest et al. (no priors)    & 0                      & 0     & 0     \\
+    \Planck likelihood only    & 0                      & 0     & 0     \\
   \end{tabular}
   \caption {Posterior distirbutions of parameters in the
-    wCDM model. Uncertainties denote symmetrized 68\%
+    wCDM model for both data sets and the \Planck likelihood only. 
+    Uncertainties denote symmetrized 68\%
     confidence intervals maginalized over other parameters.}
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/betoule_wcdm_noprior_corner.eps}
+  \includegraphics[width=0.6\linewidth]{figures/betoule_wcdm_noprior_corner.eps}
   \caption{Posterior distributions of the cosmological parameters in the
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/betoule_wcdm_withprior_corner.eps}
+  \includegraphics[width=0.6\linewidth]{figures/betoule_wcdm_withprior_corner.eps}
   \caption{Posterior distributions of the cosmological parameters in the

### 2016-05-07 Final_Project/figures/corner_code/make_corner.py

@@ -1,0 +1,33 @@
+"""
+This script takes in the chain and creates corner plots.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import corner
+plt.rc('text',usetex=True,fontsize=16)
+
+priors = False
+makecorner = True
+
+path = "prior_astr_513_chain.txt"
+
+data = np.genfromtxt(path)
+print data.shape
+
+chain = data[:,:2]
+likes = data[:,2]
+print chain.shape
+
+if makecorner:
+    fontsize = 28
+    pad = 100
+    fig = corner.corner(chain,labels=[r"$\Omega_{\rm M}$",r"$h$"],\
+                            quantiles=[0.16, 0.84],\
+                            levels=[1-np.exp(-0.5),1-np.exp(-(2.**2)/2.)],\
+                            plot_datapoints=False)
+    fig.get_axes()[2].set_ylabel(r"$h$",fontsize=fontsize)
+    fig.get_axes()[2].set_xlabel(r"$\Omega_{\rm M}$",fontsize=fontsize)
+    fig.get_axes()[3].set_xlabel(r"$h$",fontsize=fontsize)
+
+    plt.gcf().savefig("betoule_lcdm_withprior_corner.eps")
+    #plt.show()

### 2016-05-07 Final_Project/ms.tex
Added code that make corner plots. Added betoule figures for lcdm.

@@
   \caption{Posterior distributions of the cosmological parameters in the
-    analysis of the \citet{Betoule2014} data without including the
-    \Planck priors. Contours denote the
+    analysis of the \citet{betoule2014} data without including the
+    \Planck priors for a wCDM model. Contours denote the
     68\% and 95\% confidence projected areas; dashed lines in
@@
 \end{figure}
-
+%
 \begin{figure}
@@
   \caption{Posterior distributions of the cosmological parameters in the
-    analysis of the \citet{Betoule2014} data including the
-    \Planck priors. Contours denote the
+    analysis of the \citet{betoule2014} data including the
+    \Planck priors for a wCDM model. Contours denote the
     68\% and 95\% confidence projected areas; dashed lines in
@@
   \label{fig:betoule_wcdm_withprior_corner}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.6\linewidth]{figures/betoule_lcdm_noprior_corner.eps}
+  \caption{Posterior distributions of the cosmological parameters in the
+    analysis of the \citet{betoule2014} data without including the
+    \Planck priors for a $\Lambda$CDM model. Contours denote the
+    68\% and 95\% confidence projected areas; dashed lines in
+    the 1D histograms refer to the 68\% confidence intervals.}
+  \label{fig:betoule_lcdm_noprior_corner}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.6\linewidth]{figures/betoule_lcdm_withprior_corner.eps}
+  \caption{Posterior distributions of the cosmological parameters in the
+    analysis of the \citet{betoule2014} data including the
+    \Planck priors for a $\Lambda$CDM model. Contours denote the
+    68\% and 95\% confidence projected areas; dashed lines in
+    the 1D histograms refer to the 68\% confidence intervals.}
+  \label{fig:betoule_lcdm_withprior_corner}
 \end{figure}
@@
 
-\appendix
-\section{Posterior Distribution Functions Without \Planck Priors}
-\TM{Since there are potentially a bunch of corner plots, we could jam a bunch of
-  them here as well as perhaps the discussion of what happens when we remove
-  the priros. Let me know your thoughts.}
-
 \end{document}

### 2016-05-07 Final_Project/cosmosis_code/chains/wcdm/betoule_stuff/withpriors/make_corner.py
Added all the chains.

@@ -1,0 +1,35 @@
+"""
+This script takes in the chain and creates corner plots.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import corner
+plt.rc('text',usetex=True,fontsize=16)
+
+priors = True
+makecorner = True
+
+path = "prior_astr_513_chain.txt"
+
+data = np.genfromtxt(path)
+print data.shape
+
+chain = data[:,:3]
+likes = data[:,3]
+print chain.shape
+
+if makecorner:
+    fontsize = 28
+    pad = 100
+    fig = corner.corner(chain,labels=[r"$\Omega_{\rm M}$",r"$h$",r"$w$"],\
+                            quantiles=[0.16, 0.84],\
+                            levels=[1-np.exp(-0.5),1-np.exp(-(2.**2)/2.)],\
+                            plot_datapoints=False)
+    fig.get_axes()[3].set_ylabel(r"$h$",fontsize=fontsize)
+    fig.get_axes()[6].set_ylabel(r"$w$",fontsize=fontsize)
+    fig.get_axes()[6].set_xlabel(r"$\Omega_{\rm M}$",fontsize=fontsize)
+    fig.get_axes()[7].set_xlabel(r"$h$",fontsize=fontsize)
+    fig.get_axes()[8].set_xlabel(r"$w$",fontsize=fontsize)
+
+    plt.gcf().savefig("betoule_withprior_corner.eps")
+    plt.show()

### 2016-05-07 Final_Project/cosmosis_code/chains/wcdm/rest_stuff/make_corner.py

@@ -1,0 +1,35 @@
+"""
+This script takes in the chain and creates corner plots.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import corner
+plt.rc('text',usetex=True,fontsize=16)
+
+priors = False
+makecorner = True
+
+path = "noprior_astr_513_chain.txt"
+
+data = np.genfromtxt(path)
+print data.shape
+
+chain = data[:,:3]
+likes = data[:,3]
+print chain.shape
+
+if makecorner:
+    fontsize = 28
+    pad = 100
+    fig = corner.corner(chain,labels=[r"$\Omega_{\rm M}$",r"$h$",r"$w$"],\
+                            quantiles=[0.16, 0.84],\
+                            levels=[1-np.exp(-0.5),1-np.exp(-(2.**2)/2.)],\
+                            plot_datapoints=False)
+    fig.get_axes()[3].set_ylabel(r"$h$",fontsize=fontsize)
+    fig.get_axes()[6].set_ylabel(r"$w$",fontsize=fontsize)
+    fig.get_axes()[6].set_xlabel(r"$\Omega_{\rm M}$",fontsize=fontsize)
+    fig.get_axes()[7].set_xlabel(r"$h$",fontsize=fontsize)
+    fig.get_axes()[8].set_xlabel(r"$w$",fontsize=fontsize)
+
+    plt.gcf().savefig("rest_wcdm_noprior_corner.eps")
+    plt.show()

### 2016-05-07 Final_Project/ms.tex
Removed some clutter. Rearranged todo notes to be in the proper places.

@@
 \label{sec:results}
-\TM{Discuss the results and show the corner plots. Example plot included.
-  We should have one corner plot without priors and one with priors.}
 %\LC{Not sure where to put why we couldn't do the full analysis of Rest's data. 
@@
 
-\EW{Filler table for the lambda CDM fits}
 \begin{table}[h]
@@
       $\Lambda$CDM model. uncertainties denote symmetrized 68\%
-      confidence intervals maginalized over other parameters.}
+      confidence intervals maginalized over other parameters.
+      \EW{Filler table for the lambda CDM fits}}
 \end{table}
-
-\EW{Just a filler for the table for the wCDM fits}
+%
 \begin{table}[h]
@@
     Uncertainties denote symmetrized 68\%
-    confidence intervals maginalized over other parameters.}
+    confidence intervals maginalized over other parameters.
+    \EW{Just a filler for the table for the wCDM fits}}
 \end{table}
@@
   \centering
-  \includegraphics[width=0.6\linewidth]{figures/betoule_wcdm_noprior_corner.eps}
+  \includegraphics[width=0.5\linewidth]{figures/betoule_wcdm_noprior_corner.eps}
   \caption{Posterior distributions of the cosmological parameters in the
@@
   \centering
-  \includegraphics[width=0.6\linewidth]{figures/betoule_wcdm_withprior_corner.eps}
+  \includegraphics[width=0.5\linewidth]{figures/betoule_wcdm_withprior_corner.eps}
   \caption{Posterior distributions of the cosmological parameters in the

### 2016-05-07 Final_Project/ms.tex
Rearranged chain folders. Wrote the best fit plot script. Added the best fit plot for betoule to the paper.

@@
   \centering
+  \includegraphics[width=0.5\linewidth]{figures/betoule_wcdm_withprior_bestfit.eps}
+  \caption{Best fit model of the \citet{betoule2014} data including the
+    \Planck priors for a wCDM model. Blue points are the data with
+    error bars corresponding to the diagonal of the provided covariance martrix.
+    The solid black line is the most likely model from our MCMC chain.}
+  \label{fig:betoule_wcdm_noprior_bestfit}
+\end{figure}
+
+\begin{figure}
+  \centering
   \includegraphics[width=0.5\linewidth]{figures/betoule_wcdm_noprior_corner.eps}

### 2016-05-07 Final_Project/cosmosis_code/best_fit_plot.py

@@ -1,0 +1,64 @@
+"""
+This script takes in a chain and plots the most likely model.
+
+NOTE: My laptop (Tom) won't write things correctly when I force it to
+use latex. Uncomment the line plt.rc... and rerun to make the figures please.
+
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy import integrate
+#plt.rc('text',usetex=True,fontsize=24)
+
+def E_z(z,om,ode,w,ok=0):
+    return 1./np.sqrt(om*(1.0+z)**3.0+ok*(1.0+z)**2.0+ode*(1.0+z)**(3.0*(w+1.0)))
+
+betoule = True
+rest = False
+prior = "withpriors/prior"
+priorfigname = "withprior"
+modelname = "wcdm"
+
+if betoule:
+    datapath = "../Betoule2014Data/tablef1.dat"
+    covpath = "../Betoule2014Data/tablef2.dat"
+    name = "betoule"
+elif rest:
+    datapath = "../Rest2014Data/redshift_distanceMod.dat"
+    covpath = "../Rest2014Data/covarianceMatrix.dat"
+    name = "rest"
+
+figname = "%s_%s_%s_bestfit.eps"%(name,modelname,priorfigname)
+
+#Get the model
+chainpath = "chains/%s/%s_stuff/%s_astr_513_chain.txt"%(modelname,name,prior)
+chain = np.genfromtxt(chainpath)
+om,h,w,maxlike = chain[np.argmax(chain[:,-1])]
+ode = 1.0 - om
+
+data = np.genfromtxt(datapath)
+cov = np.genfromtxt(covpath)
+errs = np.sqrt(np.diag(cov)) #errorbars
+icov = np.linalg.inv(cov)
+if rest:
+    icov = np.genfromtxt(covpath) #rest matrix is mislabeled
+z,mu = data.T
+inds = np.argsort(z)
+z = z[inds]
+mu = mu[inds]
+
+#Create a model for mu(z)
+dc = np.ones_like(z)
+for i in range(len(z)):
+    dc[i] = 3000.0/h * integrate.quad(E_z,0,z[i],args=(om,ode,w))[0]
+dl = (1+z)*dc
+mu_model = 5*np.log10(dl)+25
+    
+fs = 30
+plt.errorbar(z,mu,yerr=errs,ls='',marker='.',alpha=0.5)
[…]

### 2016-05-07 Final_Project/cosmosis_code/chains/wcdm/betoule_stuff/withoutpriors/make_corner.py

@@ -1,35 +1,0 @@
-"""
-This script takes in the chain and creates corner plots.
-"""
-import numpy as np
-import matplotlib.pyplot as plt
-import corner
-plt.rc('text',usetex=True,fontsize=16)
-
-priors = False
-makecorner = True
-
-path = "noprior_astr_513_chain.txt"
-
-data = np.genfromtxt(path)
-print data.shape
-
-chain = data[:,:3]
-likes = data[:,3]
-print chain.shape
-
-if makecorner:
-    fontsize = 28
-    pad = 100
-    fig = corner.corner(chain,labels=[r"$\Omega_{\rm M}$",r"$h$",r"$w$"],\
-                            quantiles=[0.16, 0.84],\
-                            levels=[1-np.exp(-0.5),1-np.exp(-(2.**2)/2.)],\
-                            plot_datapoints=False)
-    fig.get_axes()[3].set_ylabel(r"$h$",fontsize=fontsize)
-    fig.get_axes()[6].set_ylabel(r"$w$",fontsize=fontsize)
-    fig.get_axes()[6].set_xlabel(r"$\Omega_{\rm M}$",fontsize=fontsize)
-    fig.get_axes()[7].set_xlabel(r"$h$",fontsize=fontsize)
-    fig.get_axes()[8].set_xlabel(r"$w$",fontsize=fontsize)
-
-    plt.gcf().savefig("betoule_noprior_corner.eps")
-    plt.show()

### 2016-05-07 Final_Project/cosmosis_code/chains/wcdm/betoule_stuff/withpriors/make_corner.py

@@ -1,35 +1,0 @@
-"""
-This script takes in the chain and creates corner plots.
-"""
-import numpy as np
-import matplotlib.pyplot as plt
-import corner
-plt.rc('text',usetex=True,fontsize=16)
-
-priors = True
-makecorner = True
-
-path = "prior_astr_513_chain.txt"
-
-data = np.genfromtxt(path)
-print data.shape
-
-chain = data[:,:3]
-likes = data[:,3]
-print chain.shape
-
-if makecorner:
-    fontsize = 28
-    pad = 100
-    fig = corner.corner(chain,labels=[r"$\Omega_{\rm M}$",r"$h$",r"$w$"],\
-                            quantiles=[0.16, 0.84],\
-                            levels=[1-np.exp(-0.5),1-np.exp(-(2.**2)/2.)],\
-                            plot_datapoints=False)
-    fig.get_axes()[3].set_ylabel(r"$h$",fontsize=fontsize)
-    fig.get_axes()[6].set_ylabel(r"$w$",fontsize=fontsize)
-    fig.get_axes()[6].set_xlabel(r"$\Omega_{\rm M}$",fontsize=fontsize)
-    fig.get_axes()[7].set_xlabel(r"$h$",fontsize=fontsize)
-    fig.get_axes()[8].set_xlabel(r"$w$",fontsize=fontsize)
-
-    plt.gcf().savefig("betoule_withprior_corner.eps")
-    plt.show()

### 2016-05-07 Final_Project/cosmosis_code/chains/wcdm/rest_stuff/make_corner.py

@@ -1,35 +1,0 @@
-"""
-This script takes in the chain and creates corner plots.
-"""
-import numpy as np
-import matplotlib.pyplot as plt
-import corner
-plt.rc('text',usetex=True,fontsize=16)
-
-priors = False
-makecorner = True
-
-path = "noprior_astr_513_chain.txt"
-
-data = np.genfromtxt(path)
-print data.shape
-
-chain = data[:,:3]
-likes = data[:,3]
-print chain.shape
-
-if makecorner:
-    fontsize = 28
-    pad = 100
-    fig = corner.corner(chain,labels=[r"$\Omega_{\rm M}$",r"$h$",r"$w$"],\
-                            quantiles=[0.16, 0.84],\
-                            levels=[1-np.exp(-0.5),1-np.exp(-(2.**2)/2.)],\
-                            plot_datapoints=False)
-    fig.get_axes()[3].set_ylabel(r"$h$",fontsize=fontsize)
-    fig.get_axes()[6].set_ylabel(r"$w$",fontsize=fontsize)
-    fig.get_axes()[6].set_xlabel(r"$\Omega_{\rm M}$",fontsize=fontsize)
-    fig.get_axes()[7].set_xlabel(r"$h$",fontsize=fontsize)
-    fig.get_axes()[8].set_xlabel(r"$w$",fontsize=fontsize)
-
-    plt.gcf().savefig("rest_wcdm_noprior_corner.eps")
-    plt.show()

### 2016-05-07 Final_Project/cosmosis_code/calculate_mean_and_std.py
Added a file that calculates the mean and standard deviation of the chains.

@@ -1,0 +1,23 @@
+"""
+This script calculates the mean and standard deviations
+from the chains to report in the tables.
+"""
+
+import numpy as np
+
+name = "betoule"
+prior = "withpriors/prior"
+priorfigname = "withprior"
+modelname = "wcdm"
+
+#Get the chain
+chainpath = "chains/%s/%s_stuff/%s_astr_513_chain.txt"%(modelname,name,prior)
+chain = np.genfromtxt(chainpath)
+
+mean = np.mean(chain,0)
+std = np.std(chain,0)
+
+print mean
+print std
+
+#Edit the above to print in a format suited to the paper

### 2016-05-07 Final_Project/ms.tex
best fit plot, fixed autoref

@@
 (BAO) that are gathered in \citet{planck2013}. We look at two cosmologies in our 
-analysis: the $\Lambda \textnormal{CDM}$ universe and the wCDM universe. This 
+analysis: the $\Lambda \textnormal{CDM}$ universe and the $w$CDM universe. This 
 paper is outlined as follows. In Section 2, the process and importance of {\snea} 
@@
 is our prior, and it captures any degeneracies in parameter space
-observed by the \Planck analysis, as seen in figure 3.
-%\autoref{fig:planck_corner}
-\TM{The autoref isn't working here. If someone can figure it out put it back in.}
-
+observed by the \Planck analysis, as seen in \autoref{fig:planck_corner}. 
+%
 \begin{figure}
-  \label{fig:planck_corner}
   \centering
   \includegraphics[width=0.6\linewidth]{figures/corner_example.eps}
-  \caption{\TM{THIS IS A PLACE HOLDER PLOT!}
-    Posterior distributions of the cosmological parameter
+  \caption{Posterior distributions of the cosmological parameter
     analysis using just the \Planck likelihood. Contours denote the
@@
     the 1D histograms refer to the 68\% confidence intervals.}
+  \label{fig:planck_corner}
+\TM{THIS IS A PLACE HOLDER PLOT!} 
 \end{figure}
+%
 
@@
 our analysis, \CosmoSIS provides a convenient way to link together 
-CAMB, the \Planck likelihood code (our priors), and our {\sn} likelihood 
+CAMB, the \Planck likelihood code (our priors), and our {\sna} likelihood 
 code in a single pipeline. This design choice minimized the significant
@@
   \caption{Best fit model of the \citet{betoule2014} data including the
-    \Planck priors for a wCDM model. Blue points are the data with
+    \Planck priors for a $w$CDM model. Blue points are the data with
     error bars corresponding to the diagonal of the provided covariance martrix.
     The solid black line is the most likely model from our MCMC chain.}
-  \label{fig:betoule_wcdm_noprior_bestfit}
+  \label{fig:betoule_wcdm_prior_bestfit}
 \end{figure}

### 2016-05-07 Final_Project/cosmosis_code/best_fit_plot.py

@@ -8,14 +8,14 @@ use latex. Uncomment the line plt.rc... and rerun to make the figures please.
 import numpy as np
 import matplotlib.pyplot as plt
 from scipy import integrate
-#plt.rc('text',usetex=True,fontsize=24)
+plt.rc('text',usetex=True,fontsize=24)
 
 def E_z(z,om,ode,w,ok=0):
     return 1./np.sqrt(om*(1.0+z)**3.0+ok*(1.0+z)**2.0+ode*(1.0+z)**(3.0*(w+1.0)))
 
 betoule = True
 rest = False
-prior = "withpriors/prior"
+prior = "withprior/prior"
 priorfigname = "withprior"
 modelname = "wcdm"
 

### 2016-05-08 Final_Project/ms.tex
Assocciated plots are now side by side

@@
 The correlation matrices for both of our data sets can be seen below in
-figures \autoref{fig:cmatrix_betoule} and \autoref{fig:cmatrix_rest}.
+\autoref{fig:cmatrix_both}.
 %
 \begin{figure}
-  \includegraphics[width=0.8\linewidth]{figures/Betoule_correlation.eps}
+  \includegraphics[width=0.5\linewidth]{figures/Betoule_correlation.eps}
+  \includegraphics[width=0.5\linewidth]{figures/Rest_correlation.eps}
   \caption{Correlation matrix of $\mu_0(z)$ from the
-    \citet{betoule2014} sample. Overall very few correlations exist between
-    samples, likely due to the averaging over redshift bins.}
-  \label{fig:cmatrix_betoule}
-\end{figure}
+    \citet{betoule2014} sample on the left and \citet{rest2014} sample 
+    on the right. Very few correlations exist in \citet{betoule2014} likely 
+    due to the averaging over redshift bins. Significant correlation exists in 
+    \citet{rest2014} due to dependencies on systematic variables
+    as seen in \autoref{eq:RestCovMatrix}.}
+%  \label{fig:cmatrix_betoule}
+%\end{figure}
 %
-\begin{figure}
-  \includegraphics[width=0.8\linewidth]{figures/Rest_correlation.eps}
-  \caption{Correlation matrix of $\mu_0(z)$ from the
-    \citet{rest2014} sample. Significant correlation exists between
-    different samples due to dependencies on systematic variables
-    as seen in \autoref{eq:RestCovMatrix}.}
-  \label{fig:cmatrix_rest}
+%\begin{figure}
+%  \includegraphics[width=0.8\linewidth]{figures/Rest_correlation.eps}
+%  \caption{Correlation matrix of $\mu_0(z)$ from the
+%    \citet{rest2014} sample. Significant correlation exists between
+%    different samples due to dependencies on systematic variables
+%    as seen in \autoref{eq:RestCovMatrix}.}
+  \label{fig:cmatrix_both}
 \end{figure}
@@
   \centering
-  \includegraphics[width=0.6\linewidth]{figures/corner_example.eps}
+  \includegraphics[width=0.5\linewidth]{figures/corner_example.eps}
   \caption{Posterior distributions of the cosmological parameter
@@
     \end{tabular}
-    \caption {\label{tab:lcdm}Posterior distirbutions of parameters in the
-      $\Lambda$CDM model. uncertainties denote symmetrized 68\%
-      confidence intervals maginalized over other parameters.
-      \EW{Filler table for the lambda CDM fits}}
+    \caption {\label{tab:lcdm}Posterior distributions of parameters in the
+      $\Lambda$CDM model. Uncertainties denote symmetrized 68\%
+      confidence intervals marginalized over other parameters.}
 \end{table}
@@
   \end{tabular}
-  \caption {Posterior distirbutions of parameters in the
-    wCDM model for both data sets and the \Planck likelihood only. 
+  \caption {Posterior distributions of parameters in the
+    $w$CDM model for both data sets and the \Planck likelihood only. 
     Uncertainties denote symmetrized 68\%
-    confidence intervals maginalized over other parameters.
-    \EW{Just a filler for the table for the wCDM fits}}
+    confidence intervals marginalized over other parameters.}
 \end{table}
-
+%
 \begin{figure}
@@
 \end{figure}
-
+%
 \begin{figure}
-  \centering
   \includegraphics[width=0.5\linewidth]{figures/betoule_wcdm_noprior_corner.eps}
+  \includegraphics[width=0.5\linewidth]{figures/betoule_wcdm_withprior_corner.eps}
   \caption{Posterior distributions of the cosmological parameters in the
-    analysis of the \citet{betoule2014} data without including the
-    \Planck priors for a wCDM model. Contours denote the
+    analysis of the \citet{betoule2014} data for a $w$CDM model. The left is without 
+    \Planck priors and the right is with \Planck priors. Contours denote the
     68\% and 95\% confidence projected areas; dashed lines in
     the 1D histograms refer to the 68\% confidence intervals.}
-  \label{fig:betoule_wcdm_noprior_corner}
+  \label{fig:betoule_wcdm_corner}
 \end{figure}
@@
 \begin{figure}
-  \centering
-  \includegraphics[width=0.5\linewidth]{figures/betoule_wcdm_withprior_corner.eps}
+  \includegraphics[width=0.5\linewidth]{figures/betoule_lcdm_noprior_corner.eps}
+  \includegraphics[width=0.5\linewidth]{figures/betoule_lcdm_withprior_corner.eps}
   \caption{Posterior distributions of the cosmological parameters in the
-    analysis of the \citet{betoule2014} data including the
-    \Planck priors for a wCDM model. Contours denote the
+    analysis of the \citet{betoule2014} data for a $\Lambda$CDM model. The left is without 
+    \Planck priors and the right is with \Planck priors. Contours denote the
     68\% and 95\% confidence projected areas; dashed lines in
     the 1D histograms refer to the 68\% confidence intervals.}
-  \label{fig:betoule_wcdm_withprior_corner}
+  \label{fig:betoule_lcdm_corner}
 \end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.6\linewidth]{figures/betoule_lcdm_noprior_corner.eps}
-  \caption{Posterior distributions of the cosmological parameters in the
-    analysis of the \citet{betoule2014} data without including the
-    \Planck priors for a $\Lambda$CDM model. Contours denote the
-    68\% and 95\% confidence projected areas; dashed lines in
-    the 1D histograms refer to the 68\% confidence intervals.}
-  \label{fig:betoule_lcdm_noprior_corner}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.6\linewidth]{figures/betoule_lcdm_withprior_corner.eps}
-  \caption{Posterior distributions of the cosmological parameters in the
-    analysis of the \citet{betoule2014} data including the
-    \Planck priors for a $\Lambda$CDM model. Contours denote the
-    68\% and 95\% confidence projected areas; dashed lines in
-    the 1D histograms refer to the 68\% confidence intervals.}
-  \label{fig:betoule_lcdm_withprior_corner}
-\end{figure}
+%
 

### 2016-05-08 Final_Project/ms.tex
Added plot commands and values
Added plot commands for more of the plots. Also added values in the
tables for the chains that are finished already. May rerun Rest wCDM
with priors if time, but mostly waiting for Rest LCDM with priors

@@
     ~                          & $\Omega_m$ & $h$ \\ \hline
-    Betoule et al.             & 0      & 0 \\
-    Betoule et al. (no priors) & 0      & 0 \\
+    Betoule et al.             & $0.315 \pm 0.002$      & $0.665 \pm 0.002$ \\
+    Betoule et al. (no priors) & $0.30 \pm 0.03$      & $0.701 \pm 0.007$ \\
     Rest et al.                & 0      & 0 \\
-    Rest et al. (no priors)    & 0      & 0 \\
+    Rest et al. (no priors)    & $0.5 \pm 0.3$      & $0.68 \pm 0.07$ \\
     \end{tabular}
@@
     ~                          & $\Omega_m$ & $h$ & $w$ \\ \hline
-    Betoule et al.             & 0                      & 0     & 0     \\
-    Betoule et al. (no priors) & 0                      & 0     & 0     \\
-    Rest et al.                & 0                      & 0     & 0     \\
-    Rest et al. (no priors)    & 0                      & 0     & 0     \\
-    \Planck likelihood only    & 0                      & 0     & 0     \\
+    Betoule et al.             & $0.290 \pm 0.002$                      & $0.664 \pm 0.002$     & $-0.830 \pm 0.007$     \\
+    Betoule et al. (no priors) & $0.2 \pm 0.1$                      & $0.699 \pm 0.008$     & $-0.9 \pm 0.2$     \\
+    Rest et al.                & $0.4 \pm 0.3$                      & $0.68 \pm 0.06$     & $-1.1 \pm 0.6$     \\
+    Rest et al. (no priors)    & $0.5 \pm 0.3$                      & $0.68 \pm 0.07$    & $-0.4 \pm 0.5$     \\
+    \Planck likelihood only    & $0.295 \pm 0.003$                      & $0.658 \pm 0.002$     & $-0.827 \pm 0.007$     \\
   \end{tabular}
@@
   \includegraphics[width=0.5\linewidth]{figures/betoule_wcdm_withprior_bestfit.eps}
-  \caption{Best fit model of the \citet{betoule2014} data including the
+  \includegraphics[width=0.5\linewidth]{figures/rest_wcdm_withprior_bestfit.eps}
+  \caption{Best fit model of the \citet{betoule2014} data (left) and \citet{rest2014} data (right) including the
     \Planck priors for a $w$CDM model. Blue points are the data with
@@
   \label{fig:betoule_wcdm_corner}
+\end{figure}
+%
+\begin{figure}
+  \includegraphics[width=0.5\linewidth]{figures/rest_wcdm_noprior_corner.eps}
+  \includegraphics[width=0.5\linewidth]{figures/rest_wcdm_withpriors_corner.eps}
+  \caption{Same as ~\autoref{fig:betoule_wcdm_corner} but with the \citet{rest2014} data.}
+  \label{fig:rest_wcdm_corner}
+\end{figure}
+%
+\EW{Wasn't sure where to put the \Planck only corner plot}
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\linewidth]{figures/nosn_wcdm_corner.eps}
+  \caption{Same as ~\autoref{fig:betoule_wcdm_corner} and ~\autoref{fig:rest_wcdm_corner}, but using only \Planck likelihoods without \sn{} data.}
+  \label{fig:nosnia_wcdm_corner}
 \end{figure}

### 2016-05-08 Final_Project/ms.tex
Changed little things in labels and captions, rearranged plots

@@
 \bibliographystyle{apj}
+\usepackage[labelfont=bf]{caption}
+\usepackage[labelsep=period]{caption}
 
@@
   \includegraphics[width=0.5\linewidth]{figures/Rest_correlation.eps}
-  \caption{Correlation matrix of $\mu_0(z)$ from the
-    \citet{betoule2014} sample on the left and \citet{rest2014} sample 
-    on the right. Very few correlations exist in \citet{betoule2014} likely 
-    due to the averaging over redshift bins. Significant correlation exists in 
-    \citet{rest2014} due to dependencies on systematic variables
-    as seen in \autoref{eq:RestCovMatrix}.}
-%  \label{fig:cmatrix_betoule}
-%\end{figure}
-%
-%\begin{figure}
-%  \includegraphics[width=0.8\linewidth]{figures/Rest_correlation.eps}
-%  \caption{Correlation matrix of $\mu_0(z)$ from the
-%    \citet{rest2014} sample. Significant correlation exists between
-%    different samples due to dependencies on systematic variables
-%    as seen in \autoref{eq:RestCovMatrix}.}
+  \caption{{\it Left}: Correlation matrix of $\mu_0(z)$ from the
+    \citet{betoule2014} sample. {\it Right}: Correlation matrix of $\mu_0(z)$ from 
+	the \citet{rest2014} sample. Very few correlations exist in 
+        \citet{betoule2014} likely due to the averaging over redshift bins. 
+	Significant correlation exists in \citet{rest2014} due to dependencies 
+	on systematic variables as seen in \autoref{eq:RestCovMatrix}.}
   \label{fig:cmatrix_both}
@@
 in a way that requires the user to provide bounds on the parameter space, which we
-detail for each parameter in \autoref{tab:priors}
+detail for each parameter in \autoref{tab:priors}.
 Within those bounds, we investigated using a flat prior $P_{prior}=1$ and
@@
   \centering
-  \includegraphics[width=0.5\linewidth]{figures/corner_example.eps}
+  \includegraphics[width=0.5\linewidth]{figures/nosn_wcdm_corner.eps}
   \caption{Posterior distributions of the cosmological parameter
@@
   \label{fig:planck_corner}
-\TM{THIS IS A PLACE HOLDER PLOT!} 
 \end{figure}
@@
 %from \citet{planck2013}. 
-
 \begin{table}[h]
@@
 \end{table}
+%
+\begin{figure}
+  \includegraphics[width=0.5\linewidth]{figures/betoule_lcdm_noprior_corner.eps}
+  \includegraphics[width=0.5\linewidth]{figures/betoule_lcdm_withprior_corner.eps}
+  \caption{Posterior distributions of the cosmological parameters in the
+    analysis of the \citet{betoule2014} data for a $\Lambda$CDM model. Contours denote the
+    68\% and 95\% confidence projected areas and dashed lines in
+    the 1D histograms refer to the 68\% confidence intervals. {\it Left}: without 
+    \Planck priors. {\it Right}: with \Planck priors.}
+  \label{fig:betoule_lcdm_corner}
+\end{figure}
 %
@@
 \begin{figure}
-  \centering
-  \includegraphics[width=0.5\linewidth]{figures/betoule_wcdm_withprior_bestfit.eps}
-  \includegraphics[width=0.5\linewidth]{figures/rest_wcdm_withprior_bestfit.eps}
-  \caption{Best fit model of the \citet{betoule2014} data (left) and \citet{rest2014} data (right) including the
-    \Planck priors for a $w$CDM model. Blue points are the data with
-    error bars corresponding to the diagonal of the provided covariance martrix.
-    The solid black line is the most likely model from our MCMC chain.}
-  \label{fig:betoule_wcdm_prior_bestfit}
-\end{figure}
-%
-\begin{figure}
   \includegraphics[width=0.5\linewidth]{figures/betoule_wcdm_noprior_corner.eps}
@@
   \caption{Posterior distributions of the cosmological parameters in the
-    analysis of the \citet{betoule2014} data for a $w$CDM model. The left is without 
-    \Planck priors and the right is with \Planck priors. Contours denote the
-    68\% and 95\% confidence projected areas; dashed lines in
-    the 1D histograms refer to the 68\% confidence intervals.}
+    analysis of the \citet{betoule2014} data for a $w$CDM model. Contours denote the
+    68\% and 95\% confidence projected areas and dashed lines in
+    the 1D histograms refer to the 68\% confidence intervals. {\it Left}: without 
+    \Planck priors. {\it Right}: with \Planck priors.}
   \label{fig:betoule_wcdm_corner}
@@
   \includegraphics[width=0.5\linewidth]{figures/rest_wcdm_withpriors_corner.eps}
-  \caption{Same as ~\autoref{fig:betoule_wcdm_corner} but with the \citet{rest2014} data.}
+  \caption{Posterior distributions of the cosmological parameters in the
+    analysis of the \citet{rest2014} data for a $w$CDM model. Contours denote the
+    68\% and 95\% confidence projected areas and dashed lines in
+    the 1D histograms refer to the 68\% confidence intervals. {\it Left}: without 
+    \Planck priors. {\it Right}: with \Planck priors.}
   \label{fig:rest_wcdm_corner}
@@
 %
-\EW{Wasn't sure where to put the \Planck only corner plot}
 \begin{figure}
   \centering
-  \includegraphics[width=0.5\linewidth]{figures/nosn_wcdm_corner.eps}
-  \caption{Same as ~\autoref{fig:betoule_wcdm_corner} and ~\autoref{fig:rest_wcdm_corner}, but using only \Planck likelihoods without \sn{} data.}
-  \label{fig:nosnia_wcdm_corner}
+  \includegraphics[width=0.48\linewidth]{figures/betoule_wcdm_withprior_bestfit.eps}
+  \includegraphics[width=0.48\linewidth]{figures/rest_wcdm_withprior_bestfit.eps}
+  \caption{{\it Left}: best fit model of the \citet{betoule2014} data for a $w$CDM model 
+     with \Planck priors. {\it Right}: best fit model of the \citet{rest2014} data for a $w$CDM model 
+     with \Planck priors. Blue points are the data with error bars corresponding 
+     to the diagonal of the provided covariance martrix. The solid black line is 
+     the most likely model from our MCMC chain.}
+  \label{fig:betoule_wcdm_prior_bestfit}
 \end{figure}
-%
-\begin{figure}
-  \includegraphics[width=0.5\linewidth]{figures/betoule_lcdm_noprior_corner.eps}
-  \includegraphics[width=0.5\linewidth]{figures/betoule_lcdm_withprior_corner.eps}
-  \caption{Posterior distributions of the cosmological parameters in the
-    analysis of the \citet{betoule2014} data for a $\Lambda$CDM model. The left is without 
-    \Planck priors and the right is with \Planck priors. Contours denote the
-    68\% and 95\% confidence projected areas; dashed lines in
-    the 1D histograms refer to the 68\% confidence intervals.}
-  \label{fig:betoule_lcdm_corner}
-\end{figure}
-%
 

### 2016-05-09 Final_Project/ms.tex
Wrote about half of results section

@@
 \label{sec:results}
+The most likely values and uncertainties of $\Omega_M$ and $h$ 
+in a $\Lambda$CDM model are listed in \autoref{tab:lcdm}. Included are results from
+using the \citet{betoule2014} and \citet{rest2014} data with and without the addition 
+of \Planck priors. Additionally, the 68\% and 95\% confidence contours can be seen
+in \autoref{fig:betoule_lcdm_corner} and (Rest's corner plot if available).
+
+Similary, \autoref{tab:wcdm}, \autoref{fig:betoule_wcdm_corner}, and \autoref{fig:rest_wcdm_corner}
+detail the most likely values and confidence intervals of $\Omega_M$, $h$, and $w$ 
+in a $w$CDM model. An extra entry is given in \autoref{tab:wcdm} for just our prior, the 
+\Planck likelihood. 
+\LC{Not done yet}
+
 %\LC{Not sure where to put why we couldn't do the full analysis of Rest's data. 
@@
 \begin{table}[h]
-  \label{tab:wcdm}
   \centering
@@
     confidence intervals marginalized over other parameters.}
+  \label{tab:wcdm}
 \end{table}

### 2016-05-09 Final_Project/cosmosis_code/calculate_mean_and_std.py
Added Rest LCDM with priors plots and data to tex file

@@ -5,10 +5,10 @@ from the chains to report in the tables.
 
 import numpy as np
 
-name = "betoule"
-prior = "withpriors/prior"
+name = "rest"
+prior = "withprior/prior"
 priorfigname = "withprior"
-modelname = "wcdm"
+modelname = "lcdm"
 
 #Get the chain
 chainpath = "chains/%s/%s_stuff/%s_astr_513_chain.txt"%(modelname,name,prior)

### 2016-05-09 Final_Project/cosmosis_code/calculate_mean_and_std.py
Updated chain for Rest wCDM with priors

@@ -8,7 +8,7 @@ import numpy as np
 name = "rest"
 prior = "withprior/prior"
 priorfigname = "withprior"
-modelname = "lcdm"
+modelname = "wcdm"
 
 #Get the chain
 chainpath = "chains/%s/%s_stuff/%s_astr_513_chain.txt"%(modelname,name,prior)

### 2016-05-09 Final_Project/ms.tex
Added Rest LCDM with priors plots and data to tex file
Updated best fit for Rest wCDM with priors
Updated the best fit, and also corrected the name of the plot file for
Rest LCDM with priors in the tex file. It should all run now.

@@
     Betoule et al. (no priors) & $0.30 \pm 0.03$      & $0.701 \pm 0.007$ \\
-    Rest et al.                & 0      & 0 \\
+    Rest et al.                & $0.322 \pm 0.003$      & $0.659 \pm 0.002$ \\
     Rest et al. (no priors)    & $0.5 \pm 0.3$      & $0.68 \pm 0.07$ \\
@@
 %
+\begin{figure}
+  \includegraphics[width=0.5\linewidth]{figures/rest_lcdm_noprior_corner.eps}
+  \includegraphics[width=0.5\linewidth]{figures/rest_lcdm_withprior_corner.eps}
+  \caption{Posterior distributions of the cosmological parameters in the
+    analysis of the \citet{rest2014} data for a $\Lambda$CDM model. Contours denote the
+    68\% and 95\% confidence projected areas and dashed lines in
+    the 1D histograms refer to the 68\% confidence intervals. {\it Left}: without 
+    \Planck priors. {\it Right}: with \Planck priors.
+  \TM{The figure on the right didn't appear in the repo. Please upload.}}
+  \label{fig:rest_lcdm_corner}
+\end{figure}
+%
 \begin{table}[h]
@@
     Betoule et al. (no priors) & $0.2 \pm 0.1$                      & $0.699 \pm 0.008$     & $-0.9 \pm 0.2$     \\
-    Rest et al.                & $0.4 \pm 0.3$                      & $0.68 \pm 0.06$     & $-1.1 \pm 0.6$     \\
+    Rest et al.                & $0.295 \pm 0.003$                      & $0.658 \pm 0.002$     & $-0.827 \pm 0.007$     \\
     Rest et al. (no priors)    & $0.5 \pm 0.3$                      & $0.68 \pm 0.07$    & $-0.4 \pm 0.5$     \\
@@
   \includegraphics[width=0.5\linewidth]{figures/rest_wcdm_noprior_corner.eps}
-  \includegraphics[width=0.5\linewidth]{figures/rest_wcdm_withpriors_corner.eps}
+  \includegraphics[width=0.5\linewidth]{figures/rest_wcdm_withprior_corner.eps}
   \caption{Posterior distributions of the cosmological parameters in the

### 2016-05-09 Final_Project/cosmosis_code/calculate_mean_and_std.py

@@ -5,7 +5,7 @@ from the chains to report in the tables.
 
 import numpy as np
 
-name = "rest"
+name = "betoule"
 prior = "withprior/prior"
 priorfigname = "withprior"
 modelname = "wcdm"

### 2016-05-09 Final_Project/ms.tex
Made some grammatical and reference changes
Made some grammar corrections and put in some of the references that
were missing. I’ve also added a few comments where things seem
confusing or an equation might need to be looked at.

@@
 to be pivotal in constraining cosmological parameters. In this paper, we present 
-our analysis of comparing and combining {\sna} inference of cosmological parameters 
+our analysis of comparing and combining {\sna} inferences of cosmological parameters 
 to those from other cosmological probes. In particular, we utilize the constraints 
-from the Cosmic Microwave Background (CMB) and the Baryonic Acoustic Oscillations 
+from the Cosmic Microwave Background (CMB) and Baryonic Acoustic Oscillations 
 (BAO) that are gathered in \citet{planck2013}. We look at two cosmologies in our 
@@
 is correlated to the peak luminosity, and this correlation has been calibrated using Cephied 
-variable stars located in the same galaxies as nearby \sna{}. With this information, one can 
-estimate peak luminosity of an \sna{} from its light curve, thereby providing a 
+variable stars located in the same galaxies as nearby \snea{}. With this information, one can 
+estimate peak luminosity of a \sna{} from its light curve, thereby providing a 
 standard candle with which to measure distance. They were first used in this way in the 
@@
 In 2009 the European Space Agency launched the \Planck satellite in order
-to measure the Cosmis Microwave Background with unparalleled precision.
+to measure the Cosmic Microwave Background with unparalleled precision.
 By measuring the small anisotropies in the temperature of the CMB across
@@
 for the optical depth $\tau$. We do not use this parameter
-in our {\sna} model.\citet{planck2013} report marginalized
+in our {\sna} model. \citet{planck2013} report marginalized
 posterior distributions for all parameters,
-which we report in Table~XXX.
+which we report in \autoref{tab:priors}.
+\EW{I'm assuming this should be the \Planck priors table?}
 
@@
 Telescope (HST), and several other experiments. The 613 SDSS \snea{} 
-come from the SDSS supernova survey (SDSS-SNS) of SDSS-II \citep{sako2014} 
-with redshifts of $0.05 < z < 0.4$, and 374 of them are selected from those 
-that were confirmed in the spectroscopic follow-up. The rest of the 
-\sne{} come from the compilation by \citet{conley2011}, which includes 
+came from the SDSS supernova survey (SDSS-SNS) of SDSS-II \citep{sako2014} 
+with redshifts of $0.05 < z < 0.4$, and of these, 374 were selected from confirmed \sna{} in the spectroscopic follow-up. The rest of the 
+\sne{} came from the compilation by \citet{conley2011}, which includes 
 242 spectroscopically confirmed \snea{} from SNLS at redshifts $0.2 < z < 1$, 
@@
 \label{sec:rest}
+\EW{Should we mention the issue with the covariance matrix in this section or the next? Or not at all?}
 \citet{rest2014} uses the first 1.5 year of the first telescope of 
@@
 and a compilation of low-redshift {\sna} studies. The low-redshift 
-sample consists of {\snea} collected by Calan/Tololo (Hamuy et al. 1996a), 
+sample consists of {\snea} collected by Cal\'{a}n/Tololo (Hamuy et al. 1996a), 
 CfA1 (Riess et al. 1999), CfA2 (Jha et al. 2006), CfA3 (Hicken et al. 2009a), 
@@
 curves, light curve fit information, covariance matrix, and each {\sn} 
-host information.\footnote{Data can be accessed at 
+host information\footnote{Data can be accessed at 
 \url{http://ipp0012.ifa.hawaii.edu/SN1aLC/}}. For our analysis, we 
@@
 \[
-    \mathcal{S}_{k}= 
+    \mathcal{S}_{k}(x)= 
 \begin{cases}
@@
 equation of state. In $w$CDM, constraints are placed on $\Omega_M$, $w$, 
-and $h$. In both cosmologies, the universe is spatially flat 
+and $h$. In both cosmologies, we assume the universe is spatially flat 
 ($\Omega_k = 0$) and thus $\Omega_{\Lambda}=1-\Omega_M$. 
+\EW{Added \textbf{we assume} to the last sentence because $\Lambda$CDM and $w$CDM do not necessarily assume a spatially flat universe}
 
@@
 %
+\EW{Is this equation missing the normalization constant?}
 For this work we investigate a combination of priors. \CosmoSIS is implemented
@@
 \Planck priors.
+\EW{This last sentence is confusing to me...}
 
@@
 of \Planck priors. Additionally, the 68\% and 95\% confidence contours can be seen
-in \autoref{fig:betoule_lcdm_corner} and (Rest's corner plot if available).
+in \autoref{fig:betoule_lcdm_corner} and \autoref{fig:rest_lcdm_corner}.
 
@@
     the 1D histograms refer to the 68\% confidence intervals. {\it Left}: without 
-    \Planck priors. {\it Right}: with \Planck priors.
-  \TM{The figure on the right didn't appear in the repo. Please upload.}}
+    \Planck priors. {\it Right}: with \Planck priors.}
   \label{fig:rest_lcdm_corner}

### 2016-05-09 Final_Project/ms.tex
Responded to EW's comment. Fixed a sentence about priors.

@@
 \EW{Is this equation missing the normalization constant?}
+\TM{No it is general. The equation before is normalized properly, and this one 
+  doesn't specify the prior, so it is normalized in that the prior hasn't been
+  specified yet.}
 For this work we investigate a combination of priors. \CosmoSIS is implemented
@@
 detail for each parameter in \autoref{tab:priors}.
-Within those bounds, we investigated using a flat prior $P_{prior}=1$ and
-\Planck priors.
-\EW{This last sentence is confusing to me...}
+Within those bounds, we investigated differences between using flat prior $P_{prior}=1$ on
+all parameters simultaneous in addition to \Planck priors on all parameters.
 

### 2016-05-10 Final_Project/ms.tex
More or less finished results and conclusions

@@
 posterior distributions for all parameters,
-which we report in \autoref{tab:priors}.
-\EW{I'm assuming this should be the \Planck priors table?}
+which we report in \autoref{tab:wcdm}.
 
@@
 \label{sec:rest}
-\EW{Should we mention the issue with the covariance matrix in this section or the next? Or not at all?}
 \citet{rest2014} uses the first 1.5 year of the first telescope of 
@@
 Note that the correlation of a variable with itself is 1 by definition.
-The correlation matrices for both of our data sets can be seen below in
+The correlation matrices for both of our data sets can be seen in
 \autoref{fig:cmatrix_both}.
@@
 ($\Omega_k = 0$) and thus $\Omega_{\Lambda}=1-\Omega_M$.
-\EW{Added \textbf{we assume} to the last sentence because $\Lambda$CDM and $w$CDM do not necessarily assume a spatially flat universe}
 
@@
 This is not the full picture though, since we must employ
-Bayes' Theorum in order to include any priors on our choice of 
+Bayes' Theorem in order to include any priors on our choice of 
 parameters and recover the posterior distribution function of our
@@
 %
-\EW{Is this equation missing the normalization constant?}
-\TM{No it is general. The equation before is normalized properly, and this one 
-  doesn't specify the prior, so it is normalized in that the prior hasn't been
-  specified yet.}
 For this work we investigate a combination of priors. \CosmoSIS is implemented
@@
   \centering
+  \begin{tabular}{lll}
+    Parameter & Description & Bounds \\ \hline
+    $h$ & Hubble constant$/100$ & $[0,2]$ \\
+    $\Omega_M$ & Matter fraction at $z=0$ & $[0,1]$ \\
+    $w$ & Dark energy E.O.S. & $[-2,0]$\\
+  \end{tabular}
   \caption{Bounds on cosmological parameters 
@@
   \label{tab:priors}
-  \begin{tabular}{lll}
-    Parameter & Description & Bounds \\ \hline
-    $h$ & Hubble constant$/100$ & $[0,2]$ \\
-    $\Omega_M$ & Matter fraction at $z=0$ & $[0,1]$ \\
-    $w$ & Dark energy E.O.S. & $[-2,0]$\\
-  \end{tabular}
 \end{table}
@@
 of \Planck priors. Additionally, the 68\% and 95\% confidence contours can be seen
-in \autoref{fig:betoule_lcdm_corner} and \autoref{fig:rest_lcdm_corner}.
+in \autoref{fig:betoule_lcdm_corner} and \autoref{fig:rest_lcdm_corner}. Similary, \autoref{tab:wcdm}, 
+\autoref{fig:betoule_wcdm_corner}, and \autoref{fig:rest_wcdm_corner}
+detail the most likely values and confidence intervals of $\Omega_M$, $h$, and $w$ 
+in a $w$CDM model. An extra entry is given in \autoref{tab:wcdm} for just the 
+\Planck likelihood. The best distance modulus fits to the \citet{betoule2014} and \citet{rest2014} data
+from our MCMC chains for a $w$CDM model with \Planck priors are shown in \autoref{fig:wcdm_prior_bestfit}.
 
-Similary, \autoref{tab:wcdm}, \autoref{fig:betoule_wcdm_corner}, and \autoref{fig:rest_wcdm_corner}
-detail the most likely values and confidence intervals of $\Omega_M$, $h$, and $w$ 
-in a $w$CDM model. An extra entry is given in \autoref{tab:wcdm} for just our prior, the 
-\Planck likelihood. 
-\LC{Not done yet}
-
-%\LC{Not sure where to put why we couldn't do the full analysis of Rest's data. 
-%  I'm putting it here for now.}
-%A complete analysis using the data from \citet{rest2014} is not done 
-%in our paper since there are several issues with the publicly released 
-%data. It is not apparent which uncertainties belonged to the statistical 
-%covariance matrix since only the systematic covariance matrix is clearly 
-%presented. Moreover, intrinsic dispersion of the data and photometric errors 
-%of the distances depend on nuisance parameters in addition to the covariances 
-%between the fit parameters. Fine-tuning nuisance parameters places a significant 
-%time restraint on our work and is beyond the scope of this paper. For these reasons, 
-%we only present the results for \citet{rest2014} that do not include the priors 
-%from \citet{planck2013}. 
+A complete analysis using the data from \citet{rest2014} reveals  
+several issues with their publicly released data. It is not apparent 
+which uncertainties belonged to the statistical covariance matrix since 
+only the systematic covariance matrix is clearly labeled in the released data. Moreover, intrinsic 
+dispersion of the data and photometric errors of the distances depend on nuisance 
+parameters in addition to the covariances between the fit parameters. Fine-tuning 
+nuisance parameters is omitted by us since it places a significant time restraint on our work. Due to 
+these issues, most of the values and contours calculated with the \citet{rest2014} data without 
+\Planck priors do not give reasonable results. 
+%
 \begin{table}[h]
@@
     \begin{tabular}{l|cc}
-    ~                          & $\Omega_m$ & $h$ \\ \hline
+    ~                          & $\Omega_M$ & $h$ \\ \hline
     Betoule et al.             & $0.315 \pm 0.002$      & $0.665 \pm 0.002$ \\
@@
   \begin{tabular}{l|ccc}
-    ~                          & $\Omega_m$ & $h$ & $w$ \\ \hline
+    ~                          & $\Omega_M$ & $h$ & $w$ \\ \hline
     Betoule et al.             & $0.290 \pm 0.002$                      & $0.664 \pm 0.002$     & $-0.830 \pm 0.007$     \\
@@
      the most likely model from our MCMC chain.}
-  \label{fig:betoule_wcdm_prior_bestfit}
+  \label{fig:wcdm_prior_bestfit}
 \end{figure}
@@
 \label{sec:conclusions}
-\TM{Talk about how the inclusions if priors changes the analysis.
-  State whether the priors are strong or uninformative.}
+\LC{Not sure if I should say anything more about \autoref {fig:wcdm_prior_bestfit} here.}
+As seen in \autoref{tab:lcdm} and \autoref{fig:betoule_lcdm_corner}, the addition of the \Planck priors
+to the \citet{betoule2014} data tightens the constraints on $\Omega_M$ and $h$ in a $\Lambda$CDM model. 
+Moreover, the \Planck priors vastly improve the constraint on $\Omega_M$ despite the issues in \citet{rest2014} data
+as demonstrated in \autoref{fig:rest_lcdm_corner}.
+
+Comparably, the inclusion of the \Planck priors to the \citet{betoule2014} data also tighten the 
+constraints on $\Omega_M$, $h$, and $w$ in a $w$CDM model as seen in \autoref{tab:wcdm} and  
+\autoref{fig:betoule_wcdm_corner}. This also applies to the \citet{rest2014} data as illustrated 
+in \autoref{fig:rest_wcdm_corner}, where $\Omega_M$ and $w$ are well constrained compared to the case with a flat prior. 
+The confidence contours and most likely values are nearly identical to the ones from the \Planck likelihood.
+
+We remind the readers that \citet{planck2013} accounts for nuisance and cosmological
+parameters unlike our analysis of the \citet{betoule2014} and \citet{rest2014} data, which disregards 
+nuisance parameters. Nonetheless, we conclude that the inclusion of information from other cosmological 
+probes as priors strengthens {\sna} inferences of cosmological parameters. Furthermore, \Planck priors prove
+to be informative especially in the case of problematic data.
 

### 2016-05-10 Final_Project/ms.tex
Grammatical change in ms.tex

@@
 Within those bounds, we investigated differences between using flat prior $P_{prior}=1$ on
-all parameters simultaneous in addition to \Planck priors on all parameters.
+all parameters simultaneously in addition to \Planck priors on all parameters.
 
